---
title: "af_weather"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(lubridate)
library(viridis)
library(knitr)
library(kableExtra)
library(zoo)
library(PairedData)
library(broom)
library(sf)
library(here)
library(skimr)
library(corrplot)
```

**Air Force SAR weather**

This code includes reads in and combines separate Excel worksheets into a single dataframe.  
The hourly data were produced in July 2019 by 14WS and provide WBGT, HI, RH, temperature (F), dew point (F) and UTCI (F).

*Note: For Fort Irwin, Daegu, Hawaii, Yongsan, Red Cloud files, need to insert row on spreadsheet prior to line 25 so that header starts on line 26.*



```{r read_function, message = FALSE, warning = FALSE, eval = FALSE}

# sar_file <- "D:/sar_weather/SAR_9860_Aberdeen Proving Ground.csv"  
#
sar_file_list <- list.files(path = "D:/sar_weather", pattern = "^.*\\.(csv)$") %>% 
                    paste0("D:/sar_weather/", .)

sar_file_list %>% head()


read_sar <- function(input_file) {

read_csv((input_file), col_names = TRUE, skip = 25, na = c("", "NA")) %>%
  janitor::clean_names() %>% 
  mutate(location = stringr::str_split((input_file), "9860_") %>% .[[1]] %>% .[2] %>% 
                             str_split(., ".csv") %>% .[[1]] %>% .[1],
             time = paste0(hour, ":", "00", ":", "00"), 
             date = paste(year, month, day, sep = "-"),
             dttm =  paste(date, time, sep = " ") %>% 
                 flipTime::AsDateTime()) %>% 
  dplyr::select(-c(year, month, day, hour, time, date)) %>% 
  dplyr::select(location, dttm, everything()) %>%
  rename(wbgt = wet_bulb_globe_temp,
         rh = relitive_humidity)
}

sar_df <- furrr::future_map_dfr(sar_file_list, read_sar, .progress = TRUE) 

sar_df <- sar_df %>% select(-c(x11, x12, x13))

sar_df

```

Save file and filter locations

```{r write_csv, eval = FALSE}
# save dataframe to .csv file
# save as .rds

# write_csv(sar_df, "D:/sar_df.csv", append = FALSE)
 
# sar_df <- read_csv("D:/sar_weather/sar_df.csv")

sar_df <- sar_df %>% 
  filter(location %in% c("Fort Benning", "Fort Bragg","Fort Campbell",
      "Fort Jackson", "Fort Polk", "Fort Hood",
      "Fort Stewart", "Fort Leonard Wood", "Fort Riley",
      "Fort Irwin", "Fort Bliss", "Fort Lewis",
      "Fort Sill", "Fort Carson", "Fort Gordon",
      "Fort Drum", "JB San Antonio")
  )

# write_rds(sar_df, path = "data/sar_df.rds")

```

```{r inspect_data}

sar_df <- read_rds("data/sar_df.rds")

sar_df

# Location counts
sar_df %>% 
  count(location) %>%
  head() %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", full_width = FALSE))
```

## Exploratory Plots

### Hour of day mean WBGT
```{r hour_of_day, warning = FALSE, fig.height = 12, cache = TRUE}

# Plot July hourly WBGT

sar_df %>%   
  filter(months(dttm) == "July") %>% 
  group_by(location, lubridate::year(dttm), lubridate::hour(dttm)) %>% 
  summarise(mean_wbgt = mean(wbgt, na.rm = TRUE)) %>%
  rename(year = `lubridate::year(dttm)`,
         hour = `lubridate::hour(dttm)`) %>% 
  ggplot(aes(x = hour, y = mean_wbgt, color = factor(year))) + 
    geom_point() +
    geom_smooth(aes(group = factor(year)),  size = 0.6, se = FALSE) +
    facet_wrap(~location) +
    labs(
      title = "Mean July WBGT (deg F) by Hour of Day (UTC)",
      x = "Hour of day (UTC)",
      y = "Mean WBGT (deg F)",
      caption = "Data from 14WS/CXO, produced 29 July 2019") +
    theme_bw() +
  viridis::scale_color_viridis(
      name = "Year", 
      discrete = TRUE,
      direction = -1,
      option = "inferno") 



```

### Yearly mean temperature boxplots

```{r yearly_boxplot, fig.height = 10, cache = TRUE}
# Locations 1 - 12
sar_df %>%   
  dplyr::select(location, dttm, temp_f) %>%
  mutate(location = as_factor(location),
         year = as.factor(lubridate::year(dttm))) %>%
  filter(as.numeric(location) > 2) %>%
  filter(as.numeric(location) <= 11) %>% 
  group_by(location, lubridate::year(dttm)) %>% 
  ggplot(aes(x = year, 
             y = temp_f)) +
    geom_boxplot() +
    facet_wrap(~location, ncol = 3) +
    labs(
      title = "Yearly mean temperatures (deg F)",
      x = "Year",
      y = "Temperature (deg F)",
      caption = "Data from 14WS/CXO, produced 29 July 2019") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, size = 6, hjust = 1))
```


## Compare with NLDAS-2 temperature and heat indices

```{r join_nldas}

nldas_wide <- read_rds("data/nldas_wide.rds") 

nldas_wide 


af_nldas_compare <-
  sar_df %>% 
    mutate(location = 
             str_to_lower(location, locale = "en"),
             location = str_replace_all(location, " ", "_"),
             location = recode(location, fort_benning = "fort_benning_ga", 
                                         fort_irwin = "ntc_and_fort_irwin",
                                         jb_san_antonio = "fort_sam_houston")) %>% 
    left_join(nldas_wide, by = c("location" = "installation", "dttm" = "utc_dttm")) %>% 
    dplyr::select(location, dttm, wbgt.x, heat_index.x, temp_f, wbgt_f, heat_index.y, tmp_f) %>% 
    rename(installation = location,
           utc_dttm = dttm,
           wbgt_af = wbgt.x,
           heat_index_af = heat_index.x,
           temp_af = temp_f,
           wbgt_nldas = wbgt_f,
           heat_index_nldas = heat_index.y,
           tmp_nldas = tmp_f)

af_nldas_compare

```

##  Station - NLDAS Comparison Plots (2008-2018)

### Temperature Comparison

```{r plots_temp_compare, fig.height = 20, cache = TRUE}


af_nldas_compare %>%
    dplyr::select(installation, utc_dttm, temp_af, tmp_nldas) %>% 
    na.omit() %>% 
  ggplot(aes(x = temp_af, y = tmp_nldas)) + 
      geom_point(alpha = 0.2) +
      geom_smooth(method = lm, se = FALSE) +
      facet_wrap(~installation, ncol = 3) +
      labs(
        title = "Temperature Comparison",
        x = "Air Force Station",
        y = "NLDAS") +
      theme_bw()


```

### Heat Index Comparison

```{r hi_compare, fig.height = 20, cache = TRUE}

af_nldas_compare %>%
    dplyr::select(installation, utc_dttm, heat_index_af, heat_index_nldas) %>% 
    na.omit() %>% 
  ggplot(aes(x = heat_index_af, y = heat_index_nldas)) + 
      geom_point(alpha = 0.2) +
      geom_smooth(method = lm, se = FALSE) +
      facet_wrap(~installation, ncol = 3) +
      labs(
        title = "Heat Index Comparison",
        x = "Air Force Station",
        y = "NLDAS") +
      theme_bw()

```

### WBGT Comparison

```{r wbgt_compare, fig.height = 20, cache = TRUE}


af_nldas_compare %>%
    dplyr::select(installation, utc_dttm, wbgt_af, wbgt_nldas) %>% 
    na.omit() %>% 
  ggplot(aes(x = wbgt_af, y = wbgt_nldas)) + 
      geom_point(alpha = 0.2) +
      geom_smooth(method = lm, se = FALSE) +
      facet_wrap(~installation, ncol = 3) +
      labs(
        title = "WBGT Comparison",
        x = "Air Force Station",
        y = "NLDAS") +
      theme_bw()
```


## Statistical comparison

```{r nest_nldas}

# Nest by each installation

af_nldas_nest <-
  af_nldas_compare %>% 
    nest(data = utc_dttm:tmp_nldas)

af_nldas_nest

```

### Correlation

```{r correlation}

# Correlation function

cor_fun = function(df) {
  df %>% 
    dplyr::select(2:7) %>% 
    cor(use = "pairwise.complete.obs")
}

# Map function over each installation
cor_nest <-
  map(af_nldas_nest$data, cor_fun)

# Add installation names to list elements 
names(cor_nest) <- af_nldas_nest$installation 


# Map corrplots

map2(.x = cor_nest, .y = names(cor_nest), ~ corrplot(.x, method = "number", order = "hclust", addrect = 2, title = .y, mar = c(0, 0, 1, 0)))

```


### Regression
```{r regression}

# Nest by each installation, linear models for Temp, Heat Index, WBGT

af_nldas_lm <-
  af_nldas_nest %>% 
      mutate(temp_lm = map(data, ~ lm(tmp_nldas ~ temp_af, data = .)),
             heat_index_lm = map(data, ~ lm(heat_index_nldas ~ heat_index_af, data = .)),
             wbgt_lm = map(data, ~ lm(wbgt_nldas ~ wbgt_af, data = .)),
             temp_glance = map(temp_lm, broom::glance),
             temp_tidy =   map(temp_lm, broom::tidy),
             heat_index_glance = map(heat_index_lm, broom::glance),
             heat_index_tidy =   map(heat_index_lm, broom::tidy),
             wbgt_glance = map(wbgt_lm, broom::glance),
             wbgt_tidy =   map(wbgt_lm, broom::tidy))
           
af_nldas_lm

# Linear Models Summary Table

# Temperature

af_nldas_lm %>%
  unnest(temp_tidy) %>%
  filter(term == "temp_af") %>% 
  dplyr::select(installation, estimate:p.value) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped")


af_nldas_lm %>%
  unnest(temp_glance) %>%
  dplyr::select(installation, r.squared, sigma, statistic, p.value, logLik) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped")



```


### Paired t-tests
```{r t_tests}
# https://cran.r-project.org/web/packages/BlandAltmanLeh/vignettes/Intro.html

# Temperature

af_nldas_compare %>%
    dplyr::select(installation, utc_dttm, temp_af, tmp_nldas) %>% 
    na.omit() %>% 
  group_by(installation) %>% 
  do(tidy(t.test(.$temp_af, 
                 .$tmp_nldas, 
                 mu = 0, 
                 alt = "two.sided", 
                 paired = TRUE, 
                 conf.level = 0.95))) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", full_width = FALSE))

# Heat Index

af_nldas_compare %>%
    dplyr::select(installation, utc_dttm, heat_index_af, heat_index_nldas) %>% 
    na.omit() %>% 
  group_by(installation) %>% 
  do(tidy(t.test(.$heat_index_af, 
                 .$heat_index_nldas, 
                 mu = 0, 
                 alt = "two.sided", 
                 paired = TRUE, 
                 conf.level = 0.95))) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", full_width = FALSE))


# WBGT

af_nldas_compare %>%
    dplyr::select(installation, utc_dttm, wbgt_af, wbgt_nldas) %>% 
    na.omit() %>% 
  group_by(installation) %>% 
  do(tidy(t.test(.$wbgt_af, 
                 .$wbgt_nldas, 
                 mu = 0, 
                 alt = "two.sided", 
                 paired = TRUE, 
                 conf.level = 0.95))) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", full_width = FALSE))
```


## Station Locations

```{r}

select_bases <-
  read_rds("data/select_bases.rds") %>% 
  filter(site_name %in% c("Fort Benning GA", "Fort Bragg","Fort Campbell",
      "Fort Jackson", "Fort Polk", "Fort Hood",
      "Fort Stewart", "Fort Leonard Wood", "Fort Riley",
      "NTC and Fort Irwin", "Fort Bliss", "Fort Lewis",
      "Fort Sill", "Fort Carson", "Fort Gordon",
      "Fort Drum", "Fort Sam Houston")) %>% 
  arrange(site_name)

select_bases 

# af_station_coords <- read_csv("D:sar_weather/Lat_Long Coords for Army Garrisons.csv") 

# write_rds(af_station_coords, path = "data/af_station_coords.rds")

af_station_coords <-
  read_rds("data/af_station_coords.rds") %>% 
  janitor::clean_names() %>%
  mutate(name = recode(name, `Fort Benning` = "Fort Benning GA", 
                             `Fort Irwin` = "NTC and Fort Irwin",
                             `JB San Antonio` = "Fort Sam Houston",
                              `JB Lewis McChord` = "Fort Lewis"))

af_station_sf <-         
  st_as_sf(af_station_coords, coords = c("lon", "lat"), crs = 4326) %>% 
  rename(site_name = name) %>% 
   filter(site_name %in% c("Fort Benning GA", "Fort Bragg","Fort Campbell",
      "Fort Jackson", "Fort Polk", "Fort Hood",
      "Fort Stewart", "Fort Leonard Wood", "Fort Riley",
      "NTC and Fort Irwin", "Fort Bliss", "Fort Lewis",
      "Fort Sill", "Fort Carson", "Fort Gordon",
      "Fort Drum", "Fort Sam Houston")) %>% 
  arrange(site_name)
         
af_station_sf


```


### Distance between Station and Centroid

```{r distance}

dist_from_centroid <- 
  sf::st_distance(select_bases$centroid, af_station_sf, by_element = TRUE)

af_station_distance <-
  af_station_sf %>% 
    mutate(distance = dist_from_centroid, 
           dist_km = as.numeric(distance / 1000)) %>%
  dplyr::select(-country, -distance) %>% 
  arrange(dist_km)

af_station_distance %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "condensed", full_width = FALSE))

# Summary of distances (km) from weather station to installation centroid

af_station_distance$dist_km %>% 
  summary() 

```


### Plots of Air Force weather station, base centroid, and NLDAS grid squares 

```{r join_sfs, message = FALSE, warning = FALSE, cache = TRUE}

## Load NLDAS grids

nldas_grid <- 
  here::here("data/nldas_grids", "NLDAS_Grid_Reference.shp") %>%
  sf::st_read(stringsAsFactors = FALSE) %>% 
  janitor::clean_names()

nldas_grid

## NLDAS and Installation Grid Overlap 

# Identify all NLDAS grids intersected by installation shapefiles

bases_nldas = NULL

  for (i in 1:nrow(select_bases)) {
    i_base = select_bases[i,] 
    base_nldas = st_intersection(i_base, nldas_grid) 
    bases_nldas = rbind(bases_nldas, base_nldas)
  }

bases_nldas

rm(i_base, base_nldas)

# Calculate area of installation in each NLDAS grid and weight for spatial weighted average
nldas_weights <- bases_nldas %>% 
    mutate(area = sf::st_area(.$geometry)) %>% 
  group_by(site_name) %>% 
    mutate(sum_area = sum(area),
           weight = (area / sum_area)) %>% 
    dplyr::select(-c(nldas_x, nldas_y, area, centroid, sum_area))


# NLDAS grids that intersect with bases (to map full grid squares)
intersects = NULL

  for (i in 1:nrow(select_bases)) {
    i_base = select_bases[i,] 
    base_intersect = nldas_grid %>% filter(lengths(st_intersects(., i_base)) > 0)
    intersects = rbind(intersects, base_intersect)
  }

rm(i_base, base_intersect)

# Join `site_name` to selected NLDAS grid geometries

intersects <- intersects %>% left_join(dplyr::select(as_tibble(nldas_weights), c(nldas_id, site_name)), by = "nldas_id")

intersects  


## Plots with intersecting NLDAS grids

select_bases$site_name <- purrr::set_names(select_bases$site_name)


base_grid_plot_function <- function(x) {
      ggplot() +
                 geom_sf(data = filter(select_bases, site_name == x)) +
                 geom_sf(data = filter(intersects, site_name == x), fill = NA) +
                 geom_sf(data = st_centroid(filter(select_bases, site_name == x)), shape = 3, size = 3, stroke = 1.5) +
                 geom_sf(data = filter(af_station_sf, site_name == x), shape = 4, size = 3, stroke = 1.5, color = "blue") +
                 geom_label(data = filter(intersects, site_name == x), aes(x = centerx, y = centery, label = nldas_id), size = 2) +
                  ggtitle(x) +
                  theme_bw() +
                  theme(axis.text.x = element_text(size = rel(1)),
                        axis.text.y = element_text(size = rel(1)),
                        axis.title.x = element_blank(),
                        axis.title.y = element_blank()) 
}

# base_grid_plot_function(select_bases$site_name[1])


af_station_grid_plots <-
  map(select_bases$site_name, ~base_grid_plot_function(.x))

af_station_grid_plots

```






