---
title: "case_crossover"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)
library(knitr)
library(kableExtra)
library(zoo)
library(purrr)
library(furrr)
library(table1)
library(furniture)
library(kableExtra)
library(visdat)
library(naniar)
library(modelr)
library(survival)
library(splines)
library(pspline)
library(dlnm)

```

Here we will prepare the code for case-crossover models of daily HSI outcomes based on indices of heat.  
Note: Many comments and code on this page are based on P8332 course materials. 


**Outcome**: HSI case (1 or 0)    
  *Variations*:     
    Type: hospitalization; ambulatory; reportable events    
    Diagnosis code: heat stroke + heat exhaustion; any HSI    

**Exposure**: Daily Temperature/Heat Index    
  *Variations*:  
    Averaging: max; mean; min  
    Index: temperature, heat index, WBGT 
    
**Effect modifiers**  
  Age group  
  Sex  
  Race/Ethnicity  
  Branch of Service  
  Grade/Rank group  
  BMI upon entry  
  Home of record state (region)  
  



### Construct cross-basis  

```{r dlnm, eval = FALSE}

# Construct crossbasis for index of heat (exposure)

cb_heat <- df %>% 
  dlnm::crossbasis(
  daily_index_heat,         
  lag = 7,    
  argvar = list(fun = "ns", df = 2), #  functional form of the dose-response curve 
  arglag = list(fun = "ns", df = 4))   # functional form of the lags  



summary(cb_heat)


```


## Case-crossover: DLNM Conditional logistic model 

Appropriateness: 
- very high exposure prevelance to environmental heat (near universal)
- primarily concerned as a short-term exposure with acute effects (same day/week)  

Process:  
Identify case days; select control days from days the case did not have an event; compare the exposure distribution between case and control days.

Unit of analysis is the *individual*. Each person acts as their own control.  

Control for confounding:  
Long term trends, seasonality, day of week: controlled by matching;
Weather (characterized as our exposure index);
No confounding by time invarient variables  

Assess for effect modification with interaction terms


### Create strata for case-crossover
Identify the control days for each case day and assign exposures for each day.  

```{r cc_strata, eval = FALSE}

# Option 1: function to ID control days; reference dates have the same month and wday as event 
# https://stackoverflow.com/questions/20676779/create-a-case-control-pair-for-time-stratified-case-crossover-design

control_dates_fun <- function(event_date) {
    possible_refs <- as.Date((event_date - 30) : (event_date + 30), 
                             origin = "1970-01-01")
    
    possible_refs[month(possible_refs) == lubridate::month(event_date)
                        & lubridate::wday(possible_refs) == lubridate::wday(event_date)
                        & event_date != possible_refs]
}

some_event_dates <- structure(c(12539, 12544, 12545, 12550, 
                                12563, 12567,14065), class = "Date")

some_event_dates

lapply(some_event_dates, control_dates_fun)



# Option 2: Apply function from RyanGan/case.crossover

# https://rdrr.io/github/RyanGan/case.crossover/man/casecross.html
# casecross(data, id, date, covariate = F, period = "month")
# Function creates a time-stratified case-crossover dataframe from a case-only dataframe 
# where the outcome on a particular date can be compared to referent periods on the same day of the week.
# 

#https://github.com/RyanGan/case.crossover/blob/master/R/casecross.R

casecross <- function(data, id, date, covariate=F, period = "month"){
  # if id value is given
  if(is.character(id)){
    # vector of ids
    id_vec <- data[,id]
  } else { # else if no value provided, create an id variable
    id_vec <- seq(1, nrow(data), by=1)
  }
  # vector of admit dates joined to the id vector
  event_date <- data[,date]
  id_date <- data.frame(id_vec, event_date)

  # find ref dates
  date_list <- apply(id_date, 1, function(x) {
    # output event date
    event_date <- as.Date(x[2])
    # day of week
    day_week <- wday(event_date)
    # create sequence of dates based on ref period
    date_seq <- seq.Date(floor_date(event_date, unit = period),
                         ceiling_date(event_date, unit = period), "days")
    # find dates on the same day of event date
    ref_dates <- as.character(date_seq[wday(date_seq)==day_week])

    # identifier
    identifier <- rep(x[1], length(ref_dates))
    # outcome
    outcome <- ifelse(as.Date(ref_dates) == x[2],1,0)
    id_date_vec <- cbind(identifier, ref_dates, outcome)
  })
  # bind lists together
  strat_data <- do.call(rbind, date_list)
  # remove row names
  rownames(strat_data) <- NULL
  colnames(strat_data) <- c("id", date, "outcome")
  # convert to dataframe
  strat_data <- data.frame(strat_data)
  # if covariates provided, join to the dataframe
  if(is.character(covariate)){
    cov_data <- as.data.frame(cbind(id_vec, data[,covariate]))
    # names of cov data
    colnames(cov_data) <- c("id", covariate)
    # conver identifier to character to merge
    cov_data$id <- as.character(cov_data$id)
    # merge with ts_data
    strat_data <- merge(strat_data, cov_data, by = "id")
  }
  # return dataframe
  return(strat_data)
} # end function

```

### Conditional logistic models  

```{r cc_models, eval = FALSE}

# Model (without lag)
survival::clogit(case ~ ns(daily_index_heat, df = 3) +  
               strata(id), 
               method = "efron", 
               df) 

# Model with cross-basis, no interactions
survival::clogit(case ~ cb_heat +  
               strata(id), 
               method = "efron", 
               df) 


# Models with cross-basis, interactions


survival::clogit(case ~ cb_heat +
               daily_index_heat * age +      
               strata(id), 
               method = "efron", 
               df)

survival::clogit(case ~ cb_heat +
               daily_index_heat * sex +      
               strata(id), 
               method = "efron", 
               df)

survival::clogit(case ~ cb_heat +
               daily_index_heat * race +      
               strata(id), 
               method = "efron", 
               df)
```

