---
title: "climatology"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(dygraphs)
library(xts)
library(timetk)
library(zoo)
library(forecast)
library(fable)
library(sweep)
library(viridis)
library(lubridate)
library(scales)
```


```{r read_in}
nldas_wide <- read_rds("data/nldas_wide.rds")

# Add one second to duplicate times
local_dst_adjust <-
  nldas_wide %>%
    mutate(row = dplyr::row_number()) %>% 
    tsibble::duplicates(key = installation, index = local_dttm) %>% 
    filter(local_dttm != lead(local_dttm, default = "1")) %>% 
    mutate(local_dttm = local_dttm + seconds(1))

local_dst_adjust

# Remove rows with duplicate local times and replace with updated rows (+1 second)

nldas_wide <-
  nldas_wide %>%
    mutate(row = dplyr::row_number()) %>% 
    filter(!row %in% local_dst_adjust$row) %>% 
    bind_rows(local_dst_adjust) %>% 
    arrange(local_dttm) %>% 
    dplyr::select(-row) %>% 
    dplyr::select(installation, local_dttm, tmp_f, heat_index, wbgt_f) 

```



## Construct climatology (1990-2018)

```{r climatology_function}
#Construct climatology (1990-2018)

# Create month and day columns, restrict years

nldas_wide <-
  nldas_wide %>% 
  filter(lubridate::year(local_dttm) < 2019 &
           lubridate::year(local_dttm) > 1989) %>%
    mutate(month = month(local_dttm), 
           day = day(local_dttm)) 

nldas_wide

# Nest by installation
nest_nldas <-
  nldas_wide %>% 
    nest(., data = local_dttm:day)

nest_nldas

nest_nldas %>% slice(1) %>% unnest(data)




# Define Percentile Functions 
  # https://tbradley1013.github.io/2018/10/01/calculating-quantiles-for-groups-with-dplyr-summarize-and-purrr-partial/

percentiles <- c(0.85, 0.9, 0.95)

percentile_names <- map_chr(percentiles, ~paste0(.x * 100, "%"))

percentile_funs <- map(percentiles, ~partial(quantile, probs = .x, na.rm = TRUE)) %>% 
  set_names(nm = percentile_names)

percentile_funs



# Climatology (daily) function

climatology_function <- function(df) {
  df %>% 
    group_by(month, day) %>% 
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(mean = "mean", max = "max", min = "min", sd = "sd", percentile_funs)) %>% 
    unite(mon_day, month:day, sep = "-", remove = FALSE)
}

      ## Percentiles are averaged over all daily temperatures, not daily maximums

# Map climatology function

nldas_climatology <- 
  nest_nldas %>% 
    mutate(climatology = 
      map(nest_nldas$data, climatology_function))


nldas_climatology

nldas_climatology$climatology[[1]]
```


## Anomalies (Daily)


### Daily anomaly (mean)

Mean daily temperature/index compared to the mean of the temperatures over the climatology period on the same day of year.

```{r mean_daily_anomalies}

# Single Installation

## Mean daily anomalies
nldas_climatology$data[[1]] %>% 
  mutate(date = as.Date(local_dttm)) %>% 
  group_by(date) %>% # daily aggregates
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(mean = "mean")) %>%
    ungroup() %>%
    mutate(month = month(date),
           day = day(date)) %>% 
    left_join(., nldas_climatology$climatology[[1]], by = c("month", "day")) %>% 
    mutate(tmp_f_mean_anomaly = tmp_f_mean.x - tmp_f_mean.y, 
            heat_index_mean_anomaly = heat_index_mean.x - heat_index_mean.y,
            wbgt_f_mean_anomaly = wbgt_f_mean.x - wbgt_f_mean.y) %>% 
  dplyr::select(date, tmp_f_mean_anomaly, heat_index_mean_anomaly, wbgt_f_mean_anomaly)




# Function
### Daily Mean
anomaly_daily_mean_function <- function(data, climatology) {
data %>% 
  mutate(date = as.Date(local_dttm)) %>% 
  group_by(date) %>% # daily aggregates
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(mean = "mean")) %>%
    mutate(month = month(date),
           day = day(date)) %>% 
    left_join(., climatology, by = c("month", "day")) %>% 
    mutate(tmp_f_mean_anomaly = tmp_f_mean.x - tmp_f_mean.y, 
            heat_index_mean_anomaly = heat_index_mean.x - heat_index_mean.y,
            wbgt_f_mean_anomaly = wbgt_f_mean.x - wbgt_f_mean.y) %>% 
  dplyr::select(date, tmp_f_mean_anomaly, heat_index_mean_anomaly, wbgt_f_mean_anomaly) %>% 
  group_by(year = lubridate::year(date)) %>% 
       summarise_at(
         .vars = vars(tmp_f_mean_anomaly, heat_index_mean_anomaly, wbgt_f_mean_anomaly),
            .funs = c(mean = "mean")) 
}


nldas_climatology <-
  nldas_climatology %>% 
    mutate(anomaly_daily_mean =
        map2(.x = nldas_climatology$data, .y = nldas_climatology$climatology, .f = anomaly_daily_mean_function))

nldas_climatology
nldas_climatology$anomaly_daily_mean[1]


     

```

### Daily anomaly (maximum)

Maximum daily temperature/index compared to the mean of the maximum daily temperatures on the same day of year over the climatology period.

```{r daily_anomaly_max}


left_join(
nldas_climatology$data[[1]] %>% 
  mutate(date = as.Date(local_dttm)) %>% 
  group_by(date) %>% # daily aggregates
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(max = "max")) %>%
    mutate(month = month(date),
           day = day(date)),

nldas_climatology$data[[1]] %>% 
  mutate(date = as.Date(local_dttm)) %>% 
  group_by(date) %>% # daily aggregates
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(max = "max")) %>%
    mutate(month = month(date),
           day = day(date)) %>% 
  group_by(month, day) %>% 
     summarise(mean_tmp_f_max = mean(tmp_f_max),
               mean_heat_index_max = mean(heat_index_max),
               mean_wbgt_f_max = mean(wbgt_f_max)),
by = c("month", "day")
) %>% 
  mutate(tmp_f_max_anomaly = tmp_f_max - mean_tmp_f_max, 
         heat_index_max_anomaly = heat_index_max - mean_heat_index_max,
         wbgt_f_max_anomaly = wbgt_f_max - mean_wbgt_f_max) %>% 
   group_by(year = lubridate::year(date)) %>% 
   summarise_at(
         .vars = vars(tmp_f_max_anomaly, heat_index_max_anomaly, wbgt_f_max_anomaly),
            .funs = c(mean = "mean")) 



# Function
### Daily Max
anomaly_daily_max_function <- function(data) {
  left_join(
    data %>% 
      mutate(date = as.Date(local_dttm)) %>% 
      group_by(date) %>% # daily aggregates
        dplyr::summarise_at(
              .vars = vars(tmp_f, heat_index, wbgt_f),
              .funs = c(max = "max")) %>%
        mutate(month = month(date),
               day = day(date)),
  
    data %>% 
      mutate(date = as.Date(local_dttm)) %>% 
      group_by(date) %>% # daily aggregates
        dplyr::summarise_at(
              .vars = vars(tmp_f, heat_index, wbgt_f),
              .funs = c(max = "max")) %>%
        mutate(month = month(date),
               day = day(date)) %>% 
      group_by(month, day) %>% 
         summarise(mean_tmp_f_max = mean(tmp_f_max),
                   mean_heat_index_max = mean(heat_index_max),
                   mean_wbgt_f_max = mean(wbgt_f_max)),
    by = c("month", "day")
    ) %>% 
      mutate(tmp_f_max_anomaly = tmp_f_max - mean_tmp_f_max, 
             heat_index_max_anomaly = heat_index_max - mean_heat_index_max,
             wbgt_f_max_anomaly = wbgt_f_max - mean_wbgt_f_max) %>% 
       group_by(year = lubridate::year(date)) %>% 
       summarise_at(
             .vars = vars(tmp_f_max_anomaly, heat_index_max_anomaly, wbgt_f_max_anomaly),
                .funs = c(mean = "mean")) 
  }


nldas_climatology <-
  nldas_climatology %>% 
    mutate(anomaly_daily_max =
        map(.x = nldas_climatology$data, .f = anomaly_daily_max_function))

nldas_climatology
nldas_climatology$anomaly_daily_max[1]



```




### Days mean temperature index above climate normal percentile (averaged over all hours of day)
```{r mean_anomaly_days}

# Single installation

nldas_climatology$data[[1]] %>% 
  mutate(date = as.Date(local_dttm)) %>% 
  group_by(date) %>% # daily aggregates
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(mean = "mean", max = "max", min = "min")) %>%
    ungroup() %>%
    mutate(month = month(date),
           day = day(date)) %>% 
    left_join(., nldas_climatology$climatology[[1]], by = c("month", "day")) %>% 
    group_by(year = lubridate::year(date)) %>% 
     summarise(
      days_tmp_gt85pct = sum(tmp_f_mean.x > `tmp_f_85%`), 
      days_tmp_gt90pct = sum(tmp_f_mean.x > `tmp_f_90%`),  
      days_tmp_gt95pct = sum(tmp_f_mean.x > `tmp_f_95%`),
      days_heat_index_gt85pct = sum(heat_index_mean.x > `heat_index_85%`), 
      days_heat_index_gt90pct = sum(heat_index_mean.x > `heat_index_90%`),  
      days_heat_index_gt95pct = sum(heat_index_mean.x > `heat_index_95%`), 
      days_wbgt_gt85pct = sum(wbgt_f_mean.x > `wbgt_f_85%`), 
      days_wbgt_gt90pct = sum(wbgt_f_mean.x > `wbgt_f_90%`),  
      days_wbgt_gt95pct = sum(wbgt_f_mean.x > `wbgt_f_95%`)) 



# Function

anomaly_days_function <- function(data, climatology) {
  data %>% 
    mutate(date = as.Date(local_dttm)) %>% 
    group_by(date) %>% # daily aggregates
      dplyr::summarise_at(
            .vars = vars(tmp_f, heat_index, wbgt_f),
            .funs = c(mean = "mean", max = "max", min = "min")) %>%
      ungroup() %>%
      mutate(month = month(date),
             day = day(date)) %>% 
      left_join(., climatology, by = c("month", "day")) %>% 
      group_by(year = lubridate::year(date)) %>% 
       summarise(
        days_tmp_gt85pct = sum(tmp_f_mean.x > `tmp_f_85%`), 
        days_tmp_gt90pct = sum(tmp_f_mean.x > `tmp_f_90%`),  
        days_tmp_gt95pct = sum(tmp_f_mean.x > `tmp_f_95%`),
        days_heat_index_gt85pct = sum(heat_index_mean.x > `heat_index_85%`), 
        days_heat_index_gt90pct = sum(heat_index_mean.x > `heat_index_90%`),  
        days_heat_index_gt95pct = sum(heat_index_mean.x > `heat_index_95%`), 
        days_wbgt_gt85pct = sum(wbgt_f_mean.x > `wbgt_f_85%`), 
        days_wbgt_gt90pct = sum(wbgt_f_mean.x > `wbgt_f_90%`),  
        days_wbgt_gt95pct = sum(wbgt_f_mean.x > `wbgt_f_95%`)) 
}

nldas_climatology <-
  nldas_climatology %>% 
    mutate(anomaly_days_mean =
        map2(.x = nldas_climatology$data, .y = nldas_climatology$climatology, .f = anomaly_days_function))

nldas_climatology
nldas_climatology$anomaly_days_mean[1]


```




### Days max temperature index above max daily climate normal percentile
```{r max_anomaly_days}
# Single installation

  left_join(
      nldas_climatology$data[[1]]  %>% 
      group_by(date = date(local_dttm)) %>% 
      dplyr::summarise_at(
            .vars = vars(tmp_f, heat_index, wbgt_f),
            .funs = c(max = "max")) %>% 
        mutate(month = month(date), day = day(date)), 
      nldas_climatology$data[[1]] %>% 
      group_by(date = date(local_dttm)) %>% 
      dplyr::summarise_at(
            .vars = vars(tmp_f, heat_index, wbgt_f),
            .funs = c(max = "max")) %>% 
    group_by(month = month(date), day = day(date)) %>% 
      dplyr::summarise_at(
            .vars = vars(tmp_f_max, heat_index_max, wbgt_f_max),
            .funs = percentile_funs), 
    by = c("month", "day")) %>% 
  group_by(year = lubridate::year(date)) %>% 
      dplyr::summarise(
        days_tmp_max_gt85pct = sum(tmp_f_max > `tmp_f_max_85%`), 
        days_tmp_max_gt90pct = sum(tmp_f_max > `tmp_f_max_90%`),  
        days_tmp_max_gt95pct = sum(tmp_f_max > `tmp_f_max_95%`),
        days_heat_index_max_gt85pct = sum(heat_index_max > `heat_index_max_85%`), 
        days_heat_index_max_gt90pct = sum(heat_index_max > `heat_index_max_90%`),  
        days_heat_index_max_gt95pct = sum(heat_index_max > `heat_index_max_95%`), 
        days_wbgt_max_gt85pct = sum(wbgt_f_max > `wbgt_f_max_85%`), 
        days_wbgt_max_gt90pct = sum(wbgt_f_max > `wbgt_f_max_90%`),  
        days_wbgt_max_gt95pct = sum(wbgt_f_max > `wbgt_f_max_95%`)) 
  



# Function: daily max anomalies

anomaly_days_max_function <- function(data) {
   left_join(
      data %>% 
        group_by(date = date(local_dttm)) %>% 
        dplyr::summarise_at(
              .vars = vars(tmp_f, heat_index, wbgt_f),
              .funs = c(max = "max")) %>% 
          mutate(month = month(date), day = day(date)), 
      data %>% 
      group_by(date = date(local_dttm)) %>% 
      dplyr::summarise_at(
            .vars = vars(tmp_f, heat_index, wbgt_f),
            .funs = c(max = "max")) %>% 
    group_by(month = month(date), day = day(date)) %>% 
      dplyr::summarise_at(
            .vars = vars(tmp_f_max, heat_index_max, wbgt_f_max),
            .funs = percentile_funs), 
    by = c("month", "day")) %>% 
  group_by(year = lubridate::year(date)) %>% 
      dplyr::summarise(
        days_tmp_max_gt85pct = sum(tmp_f_max > `tmp_f_max_85%`), 
        days_tmp_max_gt90pct = sum(tmp_f_max > `tmp_f_max_90%`),  
        days_tmp_max_gt95pct = sum(tmp_f_max > `tmp_f_max_95%`),
        days_heat_index_max_gt85pct = sum(heat_index_max > `heat_index_max_85%`), 
        days_heat_index_max_gt90pct = sum(heat_index_max > `heat_index_max_90%`),  
        days_heat_index_max_gt95pct = sum(heat_index_max > `heat_index_max_95%`), 
        days_wbgt_max_gt85pct = sum(wbgt_f_max > `wbgt_f_max_85%`), 
        days_wbgt_max_gt90pct = sum(wbgt_f_max > `wbgt_f_max_90%`),  
        days_wbgt_max_gt95pct = sum(wbgt_f_max > `wbgt_f_max_95%`)) 
}
  
 

nldas_climatology <-
  nldas_climatology %>% 
    mutate(anomaly_days_max =
        map(.x = nldas_climatology$data, .f = anomaly_days_max_function))


nldas_climatology
nldas_climatology$anomaly_days_max[1]





```


## Daily anomaly plots
### Days with mean temperature exceeding percentiles
```{r anomaly_day_plot}

nldas_climatology$anomaly_days_mean[[1]] %>% 
  gather(key = "metric", value = "num_days", days_tmp_gt85pct:days_wbgt_gt90pct) %>% 
  ggplot(aes(x = year, y = num_days, color = metric)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Eglin AFB Anomaly Days - mean temperatures",
      x = "Year",
      y = "Number of Days Exceeding Percentile"
    ) +
    theme(legend.position = "bottom") +
    theme_bw()

nldas_climatology$anomaly_days_mean[[1]] %>% 
  gather(key = "metric", value = "num_days", days_tmp_gt85pct:days_wbgt_gt90pct) %>% 
  ggplot(aes(x = year, y = num_days, color = metric)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(
      title = "Eglin AFB Anomaly Days - mean temperatures",
      x = "Year",
      y = "Number of Days Exceeding Percentile"
    ) +
    theme_bw()  

```


### Days with maximum temperature exceeding maximum daily percentiles

```{r anomaly_day_max_plot}

nldas_climatology$anomaly_days_max[[1]] %>% 
  gather(key = "metric", value = "num_days", days_tmp_max_gt85pct:days_wbgt_max_gt90pct) %>% 
  ggplot(aes(x = year, y = num_days, color = metric)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Eglin AFB Anomaly Days - maximum temperatures",
      x = "Year",
      y = "Number of Days Exceeding Percentile"
    ) +
    theme(legend.position = "bottom") +
    theme_bw()

nldas_climatology$anomaly_days_max[[1]] %>% 
  gather(key = "metric", value = "num_days", days_tmp_max_gt85pct:days_wbgt_max_gt90pct) %>% 
  ggplot(aes(x = year, y = num_days, color = metric)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(
      title = "Eglin AFB Anomaly Days - maximum temperatures",
      x = "Year",
      y = "Number of Days Exceeding Percentile"
    ) +
    theme_bw()  

```



## Plot climatologies

### Single plot
```{r plot_climatology_single}


nldas_climatology$climatology[[1]] %>%
    ungroup() %>% 
    dplyr::select(-c(month, day, tmp_f_sd, heat_index_sd, wbgt_f_sd)) %>%
    mutate(mon_day = paste0("2019", "-", mon_day),
           mon_day = as.Date(mon_day)) %>% 
    gather(., key = "variable", value = "value", tmp_f_mean:`wbgt_f_95%`) %>% 
    na.omit(mon_day) %>%   # Feb 29 coded as NA "leap year"
    filter(variable == c("wbgt_f_max", "wbgt_f_mean", "wbgt_f_min", "wbgt_f_85%", "wbgt_f_90%", "wbgt_f_95%")) %>% 
    ggplot(aes(x = mon_day, y = value, color = variable)) + 
    geom_smooth(alpha = 0.1) +
    geom_point() +
    scale_x_date(labels = date_format("%b-%d"),
               breaks = date_breaks("months")) +
    labs(
      title = "Eglin AFB WBGT Climatology",
      x = "Month-Day of Year",
      y = "WBGT (°F)"
    ) +
    theme(legend.position = "bottom") +
    theme_bw()
  


```


### All plots

```{r plot_climatologies_map}

# Plot function
climatology_plot <- function(df) {
  df %>%
    ungroup() %>% 
    dplyr::select(-c(month, day, tmp_f_sd, heat_index_sd, wbgt_f_sd)) %>%
    mutate(mon_day = paste0("2019", "-", mon_day),
           mon_day = as.Date(mon_day)) %>% 
    gather(., key = "variable", value = "value", tmp_f_mean:`wbgt_f_95%`) %>% 
    na.omit(mon_day) %>%   # Feb 29 coded as NA "leap year"
    filter(variable == c("wbgt_f_max", "wbgt_f_mean", "wbgt_f_min", "wbgt_f_85%", "wbgt_f_90%", "wbgt_f_95%")) %>% 
    ggplot(aes(x = mon_day, y = value, color = variable)) + 
    geom_smooth(alpha = 0.1) +
    geom_point() +
    scale_x_date(labels = date_format("%b-%d"),
               breaks = date_breaks("months")) +
    labs(
      title = "WBGT Climatology",
      x = "Month-Day of Year",
      y = "WBGT (°F)"
    ) +
    theme(legend.position = "bottom") +
    theme_bw()
}

# Map plot function over list

names(nldas_climatology$climatology) <- nldas_climatology$installation
  

map(nldas_climatology$climatology, climatology_plot)


```

