---
title: "daily_models"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)
library(knitr)
library(kableExtra)
library(zoo)
library(purrr)
library(furrr)
library(table1)
library(furniture)
library(kableExtra)
library(visdat)
library(naniar)
library(modelr)
library(survival)
library(splines)
library(pspline)
library(dlnm)

```

Here we will prepare the code for models of daily HSI outcomes based on indices of heat.  
Note: Many comments and code on this page are based on P8332 course materials. 


**Outcome**: HSI case (1 or 0)    
  *Variations*:     
    Type: hospitalization; ambulatory; reportable events    
    Diagnosis code: heat stroke + heat exhaustion; any HSI    

**Exposure**: Daily Temperature/Heat Index    
  *Variations*:  
    Averaging: max; mean; min  
    Index: temperature, heat index, WBGT 
    
**Effect modifiers**  
  Age group  
  Sex  
  Race/Ethnicity  
  Branch of Service  
  Grade/Rank group  
  BMI upon entry  
  Home of record state (region)  
  
  
** Decision points**:

Type of spline - penalized /natural
Number of lag days
Degrees of freedom (exposure, lag): Approach to evaluating

 


## Distributed lag non-linear models (DLMN)

"The modelling class of DLMs and DLNMs is applied to describe associations in which the dependency
between an exposure and an outcome is lagged in time. This lag dimension represents a new space
over which the association is defined, by describing a lag-response relationship in addition to the
usual exposure-response relationship over the space of the predictor. The dependency, characterized
in the bi-dimensional space of predictor and lag, is defined here as exposure-lag-response association" 
(dlnm version 2.3.9, 2019-03-11)



### Construct cross-basis  

"A statistical development of DLMs and DLNMs is based on the application of basis functions for
parameterizing the exposure history, namely the set of lagged exposures. Specifically, two set of basis
functions are chosen independently for modelling the exposure and lag-response relationships. These
are then combined through a special tensor product defined by bi-dimensional cross-basis functions." 
(dlnm version 2.3.9, 2019-03-11)


```{r dlnm, eval = FALSE}
# vignette("dlnmOverview")

citation("dlnm")

# Construct crossbasis for index of heat (exposure)

cb_heat <- df %>% 
  dlnm::crossbasis(
  daily_index_heat,         
  lag = 7,    
  argvar = list(fun = "ns", df = 2), #  functional form of the dose-response curve 
  arglag = list(fun = "ns", df = 4))   # functional form of the lags  



summary(cb_heat)


```


### Quasi-poisson / Time series model  

Fit with Generalized linear model (GLM) with count data  
GLM: unified framework; uses link functions *g(µ)* of the expected outcome  
Quasi-likelihood: relax variance assumption (for ovedispersion)  

GLM assumptions: 
- the outcomes (Y) are independent  
- linear relationship between the transformed outcome and the prediction  

Unit of analysis is the *day*.  

Control for confounding:  
Long term trends, seasonality, day of week, weather (characterized as our exposure)  
Adjusted for in model.  
`date` serves as a proxy for other variables that vary over time and may covary with the index of heat.


```{r time_series, eval = FALSE}

model_ts <- glm(daily_hsi ~   
                  cb_heat +    # lagged, nonlinear term for heat exposure
                  day_of_week + 
                  ns(date, df = 4 * number_years) +
                  age + sex + race + branch + grade + bmi + hor,   # same-day nonlinear term for secular trend
              family = "quasipoisson",   # distribution family
              data = df)

# Examine between 4-6 df/yr


# Extract predictions 

predictions <- dlnm::crosspred(
    cb_heat, model_ts, 
    at = 60:125, 
    bylag = 0.2,  
    cen = 0,
    cumul = TRUE)


```

### Cox Proportional-Hazards Model

```{r coxph, eval = FALSE}

# Without cross-basis
model_coxph <- survival::coxph(Surv(time_to_event, case_status) ~ 
                     ns(daily_index_heat, 3) + 
                     age + sex + race + branch + grade + bmi + hor +   
                     weights = base_population,
                  df)

# With cross-basis

model_coxph_cb <- survival::coxph(Surv(time_to_event, case_status) ~ 
                     cb_heat + 
                     age + sex + race + branch + grade + bmi + hor +   
                     weights = base_population, 
                   df)

```


## Case-crossover: DLNM Conditional logistic model 

Appropriateness: 
- very high exposure prevelance to environmental heat (near universal)
- primarily concerned as a short-term exposure with acute effects (same day/week)  

Process:  
Identify case days; select control days from days the case did not have an event; compare the exposure distribution between case and control days.

Unit of analysis is the *individual*. Each person acts as their own control.  

Control for confounding:  
Long term trends, seasonality, day of week: controlled by matching;
Weather (characterized as our exposure index);
No confounding by time invarient variables  

Assess for effect modification with interaction terms


### Create strata for case-crossover
Identify the control days for each case day and assign exposures for each day.  

```{r cc_strata, eval = FALSE}

# Option 1: function to ID control days; reference dates have the same month and wday as event 
# https://stackoverflow.com/questions/20676779/create-a-case-control-pair-for-time-stratified-case-crossover-design

control_dates_fun <- function(event_date) {
    possible_refs <- as.Date((event_date - 30) : (event_date + 30), 
                             origin = "1970-01-01")
    
    possible_refs[month(possible_refs) == lubridate::month(event_date)
                        & lubridate::wday(possible_refs) == lubridate::wday(event_date)
                        & event_date != possible_refs]
}

some_event_dates <- structure(c(12539, 12544, 12545, 12550, 
                                12563, 12567,14065), class = "Date")

some_event_dates

lapply(some_event_dates, control_dates_fun)



# Option 2: Apply function from RyanGan/case.crossover

# https://rdrr.io/github/RyanGan/case.crossover/man/casecross.html
# casecross(data, id, date, covariate = F, period = "month")
# Function creates a time-stratified case-crossover dataframe from a case-only dataframe 
# where the outcome on a particular date can be compared to referent periods on the same day of the week.
# 

#https://github.com/RyanGan/case.crossover/blob/master/R/casecross.R

casecross <- function(data, id, date, covariate=F, period = "month"){
  # if id value is given
  if(is.character(id)){
    # vector of ids
    id_vec <- data[,id]
  } else { # else if no value provided, create an id variable
    id_vec <- seq(1, nrow(data), by=1)
  }
  # vector of admit dates joined to the id vector
  event_date <- data[,date]
  id_date <- data.frame(id_vec, event_date)

  # find ref dates
  date_list <- apply(id_date, 1, function(x) {
    # output event date
    event_date <- as.Date(x[2])
    # day of week
    day_week <- wday(event_date)
    # create sequence of dates based on ref period
    date_seq <- seq.Date(floor_date(event_date, unit = period),
                         ceiling_date(event_date, unit = period), "days")
    # find dates on the same day of event date
    ref_dates <- as.character(date_seq[wday(date_seq)==day_week])

    # identifier
    identifier <- rep(x[1], length(ref_dates))
    # outcome
    outcome <- ifelse(as.Date(ref_dates) == x[2],1,0)
    id_date_vec <- cbind(identifier, ref_dates, outcome)
  })
  # bind lists together
  strat_data <- do.call(rbind, date_list)
  # remove row names
  rownames(strat_data) <- NULL
  colnames(strat_data) <- c("id", date, "outcome")
  # convert to dataframe
  strat_data <- data.frame(strat_data)
  # if covariates provided, join to the dataframe
  if(is.character(covariate)){
    cov_data <- as.data.frame(cbind(id_vec, data[,covariate]))
    # names of cov data
    colnames(cov_data) <- c("id", covariate)
    # conver identifier to character to merge
    cov_data$id <- as.character(cov_data$id)
    # merge with ts_data
    strat_data <- merge(strat_data, cov_data, by = "id")
  }
  # return dataframe
  return(strat_data)
} # end function

```


### Conditional logistic models  

```{r cc_models, eval = FALSE}

# Model (without lag)
survival::clogit(case ~ ns(daily_index_heat, df = 3) +  
               strata(id), 
               method = "efron", 
               df) 

# Model with cross-basis, no interactions
survival::clogit(case ~ cb_heat +  
               strata(id), 
               method = "efron", 
               df) 


# Models with cross-basis, interactions


survival::clogit(case ~ cb_heat +
               daily_index_heat * age +      
               strata(id), 
               method = "efron", 
               df)

survival::clogit(case ~ cb_heat +
               daily_index_heat * sex +      
               strata(id), 
               method = "efron", 
               df)

survival::clogit(case ~ cb_heat +
               daily_index_heat * race +      
               strata(id), 
               method = "efron", 
               df)
```

