---
title: "dmed_full_icd"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)
library(knitr)
library(kableExtra)
library(zoo)
library(purrr)
library(furrr)
library(table1)
library(furniture)
library(kableExtra)
library(visdat)
library(naniar)
library(modelr)
library(lme4)
library(ICC) 
library(dotwhisker)
library(broom)
library(car)
library(reshape2)
```

Annual encounter trends (All ICD codes) from DMED data.

  CONUS Army Installations:  
  
    Fort Jackson, SC     
    Fort Benning, GA    
    Fort Bragg, NC  
    Fort Campbell, KY  
    Fort Polk, LA  
    Fort Hood, TX  
    Fort Stewart, GA  
    Fort Leonard Wood, MO
    Fort Riley, KS  
    Fort Irwin, CA  
    Fort Bliss, TX  
      
Population: Active-duty Army service members  


```{r read_data}

# Outcomes
dmed_tidy_full <- 
  read_rds("data/dmed_tidy_full.rds")

dmed_tidy_full

```

## Table of HSI Ambulatory Rates
Ambulatory rates (per 1,000 persons per year) of any heat stress illness type (Army personnel)
```{r rates_table}

dmed_tidy_full <- 
  read_rds("data/dmed_tidy_full.rds")

dmed_tidy_full %>% 
  filter(type %in% "Ambulatory Data",
         category %in% "Total") %>% 
  dplyr::select(location, year, rate) %>%
  group_by(location) %>%
  pivot_wider(., names_from = year, values_from = rate) %>% 
   knitr::kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(width = "700px") 
  
dmed_tidy_full %>% 
  filter(type %in% "Hospitalizations",
         category %in% "Total") %>% 
  dplyr::select(location, year, rate) %>%
  group_by(location) %>%
  pivot_wider(., names_from = year, values_from = rate) %>% 
   knitr::kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(width = "700px") 

dmed_tidy_full %>% 
  filter(type %in% "Reportable Events",
         category %in% "Total") %>% 
  dplyr::select(location, year, rate) %>%
  group_by(location) %>%
  pivot_wider(., names_from = year, values_from = rate) %>% 
   knitr::kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(width = "700px") 
```


## Plots 

```{r plot_ambulatory_rates, cache = TRUE, warning = FALSE}
dmed_tidy_full %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Ambulatory Data",
         category == "Total") %>% 
 ggplot(aes(x = year, y = rate, color = location)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Army Ambulatory Rates - all ICD codes",
      x = "Year",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))


# By gender
dmed_tidy_full %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Ambulatory Data",
         !category == "Total") %>% 
 ggplot(aes(x = year, y = rate, color = category, shape = location)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Army Ambulatory Rates - all ICD codes",
      x = "Year",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    scale_shape_manual(values = 0:11) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))




dmed_tidy_full %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         category == "Total") %>% 
 ggplot(aes(x = year, y = rate, color = location)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Army Hospitalization Rates - all ICD codes",
      x = "Year",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))


# By gender
dmed_tidy_full %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         !category == "Total") %>% 
 ggplot(aes(x = year, y = rate, color = category, shape = location)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Army Hospitalization Rates - all ICD codes",
      x = "Year",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    scale_shape_manual(values = 0:11) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))





dmed_tidy_full %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Reportable Events",
         category == "Total") %>% 
 ggplot(aes(x = year, y = rate, color = location)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Army Reportable Event Rates - all ICD codes",
      x = "Year",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))



```


```{r join_tables}


all_icd <-
  dmed_tidy_full %>% 
    dplyr::mutate(location = dplyr::recode(location, jackson = "fort_jackson"),
           location = dplyr::recode(location, benning = "fort_benning_ga"),                 
           location = dplyr::recode(location, bragg = "fort_bragg"),
           location = dplyr::recode(location, campbell = "fort_campbell"),
           location = dplyr::recode(location, polk = "fort_polk"),
           location = dplyr::recode(location, hood = "fort_hood"),
           location = dplyr::recode(location, stewart = "fort_stewart"),
           location = dplyr::recode(location, leonardwood = "fort_leonard_wood"),
           location = dplyr::recode(location, riley = "fort_riley"),
           location = dplyr::recode(location, irwin = "ntc_and_fort_irwin"),
           location = dplyr::recode(location, bliss = "fort_bliss"),
           location = as.character(location))

#write_rds(all_icd, "data/all_icd.rds")


```


