---
title: "dmed_models_hospitalization"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)
library(knitr)
library(kableExtra)
library(zoo)
library(purrr)
library(furrr)
library(table1)
library(furniture)
library(kableExtra)
library(visdat)
library(naniar)
library(modelr)
library(lme4)
library(ICC) 
library(dotwhisker)
library(lattice)
library(ggridges)


```

Annual models from DMED data.

Scope:

  CONUS Army Installations:  
  
    Fort Jackson, SC     
    Fort Benning, GA    
    Fort Bragg, NC  
    Fort Campbell, KY  
    Fort Polk, LA  
    Fort Hood, TX  
    Fort Stewart, GA  
    Fort Leonard Wood, MO
    Fort Riley, KS  
    Fort Irwin, CA  
    Fort Bliss, TX  
      
Population: Active-duty Army service members  

**Outcome**:
  Hospitalization (In-patient)  
  Annual Rate of Heat Stress Illness (any type, primary diagnosis)  
  1997 - 2018
  
**Exposure indices**:  
*“Absolute” indices*  
  Annual mean (full-year): temperature, heat index, WBGT  
  Annual heat risk days / hours - Heat index above 80 / 90 / 103 / 125 °F - WBGT above 82 / 85 / 88 / 90 °F  
  
*“Relative” indices* (averaged over full-year and heat season months) 
Annual mean daily anomaly: temperature, heat index, WBGT  
Annual maximum daily anomaly: temperature, heat index, WBGT  
Days mean temperature index above daily climate normal percentile (averaged over all hours of day)  
  - temperature, heat index, WBGT above 85th / 90th / 95th percentiles  
Days maximum temperature index above daily climate normal maximum percentile  
  - temperature, heat index, WBGT above 85th / 90th / 95th percentiles  
Days mean temperature index above Standard Deviation(s) of mean daily temperature climate normal  
  - temperature, heat index, WBGT above 1 or 2 standard deviations of daily normal  
Days maximum temperature index above Standard Deviation(s) of max daily temperature climate normal  
  - temperature, heat index, WBGT above 1 or 2 standard deviations of maximum daily normal  

```{r read_data}

# Outcomes
dmed_tidy <- 
  read_rds("data/dmed_tidy.rds")

dmed_tidy

# Exposures
annual_tables_list <-
  read_rds("data/annual_tables_list.rds")

annual_tables_list

```

## Table of HSI Hospitalization Rates
Hospitalization rates (per 1,000 persons per year) of any heat stress illness type (Army personnel)
```{r rates_table}

dmed_tidy <- 
  read_rds("data/dmed_tidy.rds")

dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
  dplyr::select(location, year, rate) %>% 
  pivot_wider(., names_from = year, values_from = rate) %>% 
   knitr::kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(width = "700px") 
  

```

## Plots of HSI Hospitalization Rates

```{r plot_hospitalization_rates, cache = TRUE, warning = FALSE}
dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
 ggplot(aes(x = year, y = rate, color = location)) +
    geom_point() +
    geom_line() +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates",
      x = "Year",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))


dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
 ggplot(aes(x = fct_reorder(location, rate), y = rate)) +
    geom_boxplot() +
    coord_flip() +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates, 1990-2018",
      x = "Installation",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() 


dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
 ggplot(aes(x = rate, y = fct_reorder(location,rate))) +
    ggridges::geom_density_ridges(scale = 0.75) +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates, 1990-2018",
      x = "Installation",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() 




```





```{r join_tables}

index_long <- 
  annual_tables_list %>% 
    filter(installation %in% c("fort_benning_ga", "fort_bragg", "fort_campbell", "fort_jackson",
                               "fort_polk", "fort_hood", "fort_stewart", "fort_leonard_wood", "fort_riley",
                               "ntc_and_fort_irwin", "fort_bliss")) %>% 
    unnest(data) %>% 
    pivot_longer(., cols = `1990`:`2019`, names_to = "year", values_to = "value") %>% 
    mutate(year = as.integer(year)) 


hsi_rates <-
  dmed_tidy %>% 
    filter(location %in% c("benning", "bragg", "campbell", "jackson",
                           "polk", "hood", "stewart", "leonardwood", 
                           "riley", "irwin", "bliss"),
           type == "Hospitalizations",
           hsi == "all",
           strata == "gender",
           category == "Total") %>% 
    dplyr::select(location, year, rate) %>% 
    mutate(location = recode(location, jackson = "fort_jackson"),
           location = recode(location, benning = "fort_benning_ga"),                 
           location = recode(location, bragg = "fort_bragg"),
           location = recode(location, campbell = "fort_campbell"),
           location = recode(location, polk = "fort_polk"),
           location = recode(location, hood = "fort_hood"),
           location = recode(location, stewart = "fort_stewart"),
           location = recode(location, leonardwood = "fort_leonard_wood"),
           location = recode(location, riley = "fort_riley"),
           location = recode(location, irwin = "ntc_and_fort_irwin"),
           location = recode(location, bliss = "fort_bliss"),
           location = as.character(location))

hsi_rates

# Join table

joined_rate <-
  index_long %>% 
    left_join(hsi_rates, by = c("installation" = "location", "year" = "year")) %>% 
    drop_na(rate)

joined_rate
          
 
          
```

## Scatterplots of Temperature/Heat Index and HSI Rates

Each point represents a year from 1997 - 2018.  
The back line is a linear regression and the blue curve is a `loess` smoothed conditional means curve.

```{r scatterplots, fig.height = 14, fig.width = 10, warning = FALSE, cache = TRUE}

joined_rate %>%
    filter(installation == "fort_benning_ga") %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates and Annual Temperature/Heat Indices \n Fort Benning, GA",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)",
      caption = "Primary diagnosis"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

 joined_rate %>%
    filter(installation == "fort_jackson") %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates and Annual Temperature/Heat Indices \n Fort Jackson, SC",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)",
      caption = "Primary diagnosis"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
 joined_rate %>%
    filter(installation == "fort_campbell") %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates and Annual Temperature/Heat Indices \n Fort Campbell, KY",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)",
      caption = "Primary diagnosis"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
    
    
  

```

## Linear models

```{r linear_models}

# Linear regression function
annual_lm = function(df) {
  df %>% 
    lm(rate ~ value, data = .)
}


# Include year in model
annual_year_lm = function(df) {
  df %>% 
    lm(rate ~ value + year, data = .)
}





# Nest by each installation - index pair

annual_hsi_nest <-
  joined_rate %>% 
    nest(data = year:rate) %>% 
    mutate(linear_models = map(data, annual_lm),
           glance = map(linear_models, broom::glance),
           tidy =   map(linear_models, broom::tidy))
           
  

annual_hsi_year_nest <-
  joined_rate %>% 
    nest(data = year:rate) %>% 
    mutate(linear_models = map(data, annual_year_lm),
           glance = map(linear_models, broom::glance),
           tidy =   map(linear_models, broom::tidy),
           augment = map(linear_models, broom::augment))  

annual_hsi_nest
annual_hsi_year_nest


```


## Linear Models Summary Table
All models, sorted by R^2^

```{r lm_models_sort}
# Model summaries

annual_hsi_nest %>%
  unnest(glance) %>% 
  dplyr::select(installation, index, r.squared:df.residual) %>%
  arrange(desc(r.squared)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")


annual_hsi_nest %>%
  unnest(tidy) %>%
  dplyr::select(installation, index, term:p.value) %>%
  filter(term %in% "value") %>% 
  arrange(p.value) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")

```


## Linear Models Summary, by Installation
```{r lm_models_grouped_sort}
# Model Summary Tables by Installation

annual_hsi_nest %>%
  unnest(glance) %>% 
  dplyr::select(installation, index, r.squared:df.residual) %>%
  group_by(installation, add = TRUE) %>% 
  arrange(desc(r.squared), .by_group = TRUE) %>% 
  group_split() %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "300px")


annual_hsi_nest %>%
  unnest(tidy) %>% 
  dplyr::select(installation, index, term:p.value) %>%
  filter(term %in% "value") %>% 
  group_by(installation, add = TRUE) %>% 
  arrange(p.value) %>% 
  group_split() %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "300px")
```

Note: NA's are present for models where the conditions were not met (e.g. no days with HI above 125°F, no days with mean temp above 95th percentile). 



### Include `year` in model.

```{r  lm_models_year}
annual_hsi_year_nest %>%
  unnest(glance) %>%
  dplyr::select(installation, index, r.squared:df.residual) %>% 
  arrange(desc(r.squared)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")


annual_hsi_year_nest %>%
  unnest(tidy) %>%
  dplyr::select(installation, index, term:p.value) %>%
  filter(term %in% "value") %>% 
  arrange(p.value) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")


```






## Full / Pooled Model

Hospitalization rates, pooled: unweighted by base populations, no indicator variables for installation

```{r full_model, warning = FALSE}

joined_rate %>% 
  filter(index %in% c("days_hi_gt80", "days_wbgt_gt82", "hours_wbgt_gt82",
                      "days_wbgt_gt85", "days_tmp_gt85pct", "days_wbgt_max_gt85pct",
                      "days_wbgt_max_gt1sd", "days_heat_index_gt85pct", 
                      "days_wbgt_max_gt90pct", "days_tmp_gt90pct_may_sep")) %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates and Annual Temperature/Heat Indices \n 
      Pooled across 11 installations (unweighted)",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 10)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

# Nest by index of heat

index_nest <-
  joined_rate %>%
    dplyr::select(index, everything()) %>% 
    nest(data = installation:rate) %>% 
    mutate(linear_models = map(data, annual_lm),
           glance = map(linear_models, broom::glance),
           tidy =   map(linear_models, broom::tidy))
           
 
index_nest %>%
  unnest(glance) %>% 
  dplyr::select(index, r.squared:df.residual) %>%
  arrange(desc(r.squared)) %>% 
   knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 8) %>% 
  scroll_box(height = "700px")

index_nest %>%
  unnest(tidy) %>% 
  filter(term %in% "value") %>%
  dplyr::select(index, term:p.value) %>%
  arrange(desc(estimate)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")


```


## Pooled linear models, weighted by population


```{r pooled_weighted}

# Join Army population by installation and year

dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
  dplyr::select(location, year, population) %>% 
  ggplot(aes(x = year, y = population, color = location)) +
    geom_point() +
    geom_line() + 
    labs(
      title = "Army Installation Population",
      x = "Year",
      y = "Population"
    ) +
  theme_bw()


dmed_population <-
  dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Hospitalizations",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
  dplyr::select(location, year, population) %>% 
    mutate(location = recode(location, jackson = "fort_jackson"),
           location = recode(location, benning = "fort_benning_ga"),                 
           location = recode(location, bragg = "fort_bragg"),
           location = recode(location, campbell = "fort_campbell"),
           location = recode(location, polk = "fort_polk"),
           location = recode(location, hood = "fort_hood"),
           location = recode(location, stewart = "fort_stewart"),
           location = recode(location, leonardwood = "fort_leonard_wood"),
           location = recode(location, riley = "fort_riley"),
           location = recode(location, irwin = "ntc_and_fort_irwin"),
           location = recode(location, bliss = "fort_bliss"),
           location = as.character(location))





joined_rate %>%
  filter(index %in% c("days_hi_gt80", "days_wbgt_gt82", "hours_wbgt_gt82",
                      "days_wbgt_gt85", "days_tmp_gt85pct", "days_wbgt_max_gt85pct",
                      "days_wbgt_max_gt1sd", "days_heat_index_gt85pct", 
                      "days_wbgt_max_gt90pct", "days_tmp_gt90pct_may_sep")) %>%
  left_join(dmed_population, by = c("installation" = "location", "year")) %>% 
    ggplot(aes(x = value, y = rate, size = population)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = lm, se = TRUE, size = 0.5, color = "black", mapping = aes(weight = population)) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates and Annual Temperature/Heat Indices \n 
      Pooled across 11 installations (weighted by populatio)",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 10)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))


  

# Weighted linear regression function

weighted_lm = function(df) {
  df %>% 
    lm(rate ~ value, weights = population, data = .)
}




weighted_nest <-
joined_rate %>% 
  left_join(dmed_population, by = c("installation" = "location", "year")) %>% 
  dplyr::select(index, everything()) %>% 
    nest(data = installation:population) %>% 
    mutate(linear_models = map(data, weighted_lm),
           glance = map(linear_models, broom::glance),
           tidy =   map(linear_models, broom::tidy))
  
weighted_nest %>%
  unnest(glance) %>% 
  dplyr::select(index, r.squared:df.residual) %>%
  arrange(desc(r.squared)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")
  
  
weighted_nest %>%
  unnest(tidy) %>% 
  filter(term %in% "value") %>%
  dplyr::select(index, term:p.value) %>%
  arrange(desc(estimate)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")   



```


## Intra-class correlation (ICC)

How much of the variability is explained by the variability across installations.
```{r icc}

#  Estimate the intra-class correlation of the exposure
# Example: `days_wbgt_gt82`

joined_rate %>% 
  filter(index %in% "days_wbgt_gt82") %>% 
  ICC::ICCest(x = installation, y = value, alpha = 0.05, CI.type = c("THD", "Smith"))




#  Estimate the intra-class correlation of the outcome

joined_rate %>% 
  filter(index %in% "days_wbgt_gt82") %>% 
  ICC::ICCest(x = installation, y = rate, alpha = 0.05, CI.type = c("THD", "Smith"))

```

Exposure index: 
  ICC > 0.75  
  *"excellent reproducibility"*  
  
Outcome (Hospitalization rate): 
  ICC approx. 0.50  
  *"fair to good reproducibility"*  
  
## Fixed intercepts

Within subject effect estimate: models estimate intercept for each installation.

```{r fixed_intercepts, warning = FALSE}


fixed_intercept_lm = function(df) {
  df %>% 
    lm(rate ~ value + year + installation, data = .)
}

# year as a confounder, installation as fixed intercept


fixed_intercept_nest <-
  joined_rate %>%
    left_join(dmed_population, by = c("installation" = "location", "year")) %>% 
    dplyr::select(index, everything()) %>% 
      nest(data = installation:population) %>% 
      mutate(linear_models = map(data, fixed_intercept_lm),
             glance = map(linear_models, broom::glance),
             tidy =   map(linear_models, broom::tidy),
             augment =  map(linear_models, broom::augment))

# Model summary tables
  # By model / annual index of heat 

fixed_intercept_nest %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:p.value) %>% 
  group_split(index) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "200px")  

  # Arranged by effect estimate
fixed_intercept_nest %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:p.value) %>% 
  filter(term %in% "value") %>% 
  arrange(desc(estimate)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px")   



# Model fits
fixed_intercept_nest %>% 
  unnest(glance, .drop = TRUE) %>% 
   dplyr::select(index, r.squared:df.residual) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px")  



# Boxplots of residuals by installation
map(fixed_intercept_nest$augment[1:5], ~
      ggplot(.x, aes(x = installation, y = .resid))+
          geom_boxplot() + 
          coord_flip())

# Coefficient plot
map(fixed_intercept_nest$linear_models[1:5], ~dotwhisker::dwplot(.x) + geom_vline(xintercept = 0, lty = 2))


```

Note: without including `year`, R^2 approx. 52-55%; with `year` R^2 62-63%

With fixed intercepts added, there is very little difference in model fits based on index selection.



## Random intercepts and slopes

Assume independence across subjects *i* but not within subjects *j*. Account for within subject correlation with a random term.  
  
`(1|installation)` designates intercepts only by random factor  
`(value|installation)`designates slopes only by random factor  
`(1 + value|installation)` designates intercepts and slopes by random factor  



```{r random_intercept_slopes, message = FALSE, warning = FALSE, cache = TRUE}

# https://stats.stackexchange.com/questions/13166/rs-lmer-cheat-sheet

# Examine evidence of random intercepts and/or slopes

# Adjusted x axis scales
joined_rate %>%
   filter(index %in% "days_wbgt_gt82") %>% 
  ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ installation, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates and Annual Days with WBGT > 82 deg F",
      x = "Days WBGT > 82",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() 

# Fixed x axis scales
joined_rate %>%
   filter(index %in% "days_wbgt_gt82") %>% 
  ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ installation) +
    labs(
      title = "Army Heat Stress Illness Hospitalization Rates and Annual Days with WBGT > 82 deg F",
      x = "Days WBGT > 82",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() 


# Function for linear mixed model (random interceps and slopes)

random_slopes_intercepts_lmer = function(df) {
  df %>% 
    lmer(rate ~ value + year + (1 + value|installation), data = .)
}

# Map regression function over each index

mixed_nest <-
  joined_rate %>%
  left_join(dmed_population, by = c("installation" = "location", "year")) %>% 
    dplyr::select(index, everything()) %>% 
      nest(data = installation:population) %>% 
      mutate(linear_models = map(data, random_slopes_intercepts_lmer),
             glance = map(linear_models, broom::glance),
             tidy =   map(linear_models, broom::tidy),
             augment = map(linear_models, broom::augment))

# returns `boundary (singular) fit`: indicates models are overfitted – random effects structure is too complex to be supported by the data


mixed_nest %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px")  

  # Arranged by effect estimate
mixed_nest %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>% 
  filter(term %in% "value") %>% 
  arrange(desc(estimate)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px")   


  # Model fit
mixed_nest %>% 
  unnest(glance, .drop = TRUE) %>% 
   dplyr::select(index, sigma:df.residual) %>%
    arrange(logLik) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px") 

# https://bbolker.github.io/morelia_2018/notes/mixedlab.html
# Boxplots of residuals by installation
map(mixed_nest$augment[1:5], ~
      ggplot(.x, aes(x = installation, y = .resid))+
          geom_boxplot() + 
          coord_flip())

# Plot random effects to look for outliers
map(mixed_nest$linear_models[1:5], ~lattice::dotplot(ranef(.x, condVar = TRUE)))


# Coefficient plot
map(mixed_nest$linear_models[1:5], ~dotwhisker::dwplot(.x) + geom_vline(xintercept = 0, lty = 2))







```

## Random intercept only models

```{r random_intercept, message = FALSE, warning = FALSE, cache = TRUE}

# Function for linear mixed model (random interceps only)

random_intercepts_lmer = function(df) {
  df %>% 
    lmer(rate ~ value + year + (1|installation), data = .)
}



random_intercepts_nest <-
  joined_rate %>%
  left_join(dmed_population, by = c("installation" = "location", "year")) %>% 
    dplyr::select(index, everything()) %>% 
      nest(data = installation:population) %>% 
      mutate(linear_models = map(data, random_intercepts_lmer),
             glance = map(linear_models, broom::glance),
             tidy =   map(linear_models, broom::tidy),
             augment = map(linear_models, broom::augment))

# 2 warnings for `fixed-effect model matrix is rank deficient so dropping 1 column / coefficient`


# Model summaries

random_intercepts_nest %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px")  

  # Arranged by effect estimate
random_intercepts_nest %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>% 
  filter(term %in% "value") %>% 
  arrange(desc(estimate)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px")   


  # Model fit
random_intercepts_nest %>% 
  unnest(glance, .drop = TRUE) %>% 
   dplyr::select(index, sigma:df.residual) %>%
    arrange(logLik) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "400px")  





# Boxplots of residuals by installation
map(random_intercepts_nest$augment[1:5], ~
      ggplot(.x, aes(x = installation, y = .resid))+
          geom_boxplot() + 
          coord_flip())

# Plot random effects to look for outliers
map(random_intercepts_nest$linear_models[1:5], ~lattice::dotplot(ranef(.x, condVar = TRUE)))


# Coefficient plot
map(random_intercepts_nest$linear_models[1:5], ~dotwhisker::dwplot(.x) + geom_vline(xintercept = 0, lty = 2))
```

