---
title: "dmed_models"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)
library(knitr)
library(kableExtra)
library(zoo)
library(purrr)
library(furrr)
library(table1)
library(furniture)
library(kableExtra)
library(visdat)
library(naniar)
library(modelr)

```

Annual models from DMED data.


Scope:

  CONUS Army Installations:  
  
    Fort Jackson, SC     
    Fort Benning, GA    
    Fort Bragg, NC  
    Fort Campbell, KY  
    Fort Polk, LA  
    Fort Hood, TX  
    Fort Stewart, GA  
    Fort Leonard Wood, MO
    Fort Riley, KS  
    Fort Irwin, CA  
    Fort Bliss, TX  
      
Population: Active-duty Army service members  

**Outcome**:
  Ambulatory (Out-patient)  
  Annual Rate of Heat Stress Illness (any type, primary diagnosis)  
  1997 - 2018
  
**Exposure indices**:  
*“Absolute” indices*  
  Annual mean (full-year): temperature, heat index, WBGT  
  Annual heat risk days / hours - Heat index above 80 / 90 / 103 / 125 °F - WBGT above 82 / 85 / 88 / 90 °F  
  
*“Relative” indices* (averaged over full-year and heat season months) 
Annual mean daily anomaly: temperature, heat index, WBGT  
Annual maximum daily anomaly: temperature, heat index, WBGT  
Days mean temperature index above daily climate normal percentile (averaged over all hours of day)  
  - temperature, heat index, WBGT above 85th / 90th / 95th percentiles  
Days maximum temperature index above daily climate normal maximum percentile  
  - temperature, heat index, WBGT above 85th / 90th / 95th percentiles  
Days mean temperature index above Standard Deviation(s) of mean daily temperature climate normal  
  - temperature, heat index, WBGT above 1 or 2 standard deviations of daily normal  
Days maximum temperature index above Standard Deviation(s) of max daily temperature climate normal  
  - temperature, heat index, WBGT above 1 or 2 standard deviations of maximum daily normal  

```{r read_data}

# Outcomes
dmed_tidy <- 
  read_rds("data/dmed_tidy.rds")

dmed_tidy

# Exposures
annual_tables_list <-
  read_rds("data/annual_tables_list.rds")

annual_tables_list

```

## Table of HSI Ambulatory Rates
Ambulatory rates (per 1,000 persons per year) of any heat stress illness type (Army personnel)
```{r rates_table}

dmed_tidy <- 
  read_rds("data/dmed_tidy.rds")

dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Ambulatory Data",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
  dplyr::select(location, year, rate) %>% 
  pivot_wider(., names_from = year, values_from = rate) %>% 
   knitr::kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(width = "700px") 
  

```



```{r join_tables}

index_long <- 
  annual_tables_list %>% 
    filter(installation %in% c("fort_benning_ga", "fort_bragg", "fort_campbell", "fort_jackson",
                               "fort_polk", "fort_hood", "fort_stewart", "fort_leonard_wood", "fort_riley",
                               "ntc_and_fort_irwin", "fort_bliss")) %>% 
    unnest(data) %>% 
    pivot_longer(., cols = `1990`:`2019`, names_to = "year", values_to = "value") %>% 
    mutate(year = as.integer(year)) 


hsi_rates <-
  dmed_tidy %>% 
    filter(location %in% c("benning", "bragg", "campbell", "jackson",
                           "polk", "hood", "stewart", "leonardwood", 
                           "riley", "irwin", "bliss"),
           type == "Ambulatory Data",
           hsi == "all",
           strata == "gender",
           category == "Total") %>% 
    dplyr::select(location, year, rate) %>% 
    mutate(location = recode(location, jackson = "fort_jackson"),
           location = recode(location, benning = "fort_benning_ga"),                 
           location = recode(location, bragg = "fort_bragg"),
           location = recode(location, campbell = "fort_campbell"),
           location = recode(location, polk = "fort_polk"),
           location = recode(location, hood = "fort_hood"),
           location = recode(location, stewart = "fort_stewart"),
           location = recode(location, leonardwood = "fort_leonard_wood"),
           location = recode(location, riley = "fort_riley"),
           location = recode(location, irwin = "ntc_and_fort_irwin"),
           location = recode(location, bliss = "fort_bliss"),
           location = as.character(location))

hsi_rates

# Join table

joined_rate <-
  index_long %>% 
    left_join(hsi_rates, by = c("installation" = "location", "year" = "year")) %>% 
    drop_na(rate)

joined_rate
          
 
          
```

## Scatterplots of Temperature/Heat Index and HSI Rates

Each point represents a year from 1997 - 2018.  
The back line is a linear regression and the blue curve is a `loess` smoothed conditional means curve.

```{r scatterplots, fig.height = 14, fig.width = 10, warning = FALSE}

joined_rate %>%
    filter(installation == "fort_benning_ga") %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Ambulatory Rates and Annual Temperature/Heat Indices \n Fort Benning, GA",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)",
      caption = "Primary diagnosis"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

 joined_rate %>%
    filter(installation == "fort_jackson") %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Ambulatory Rates and Annual Temperature/Heat Indices \n Fort Jackson, SC",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)",
      caption = "Primary diagnosis"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
 joined_rate %>%
    filter(installation == "fort_campbell") %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Ambulatory Rates and Annual Temperature/Heat Indices \n Fort Campbell, KY",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)",
      caption = "Primary diagnosis"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 6)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
    
    
  

```

## Linear models

```{r linear_models}

# Linear regression function
annual_lm = function(df) {
  df %>% 
    lm(rate ~ value, data = .)
}

# Nest by each installation - index pair

annual_hsi_nest <-
  joined_rate %>% 
    nest(data = year:rate) %>% 
    mutate(linear_models = map(data, annual_lm),
           glance = map(linear_models, broom::glance),
           tidy =   map(linear_models, broom::glance))
           
  

annual_hsi_nest



```


## Linear Models Summary Table
All models, sorted by R^2^

```{r lm_models_sort}
# Model summaries

annual_hsi_nest %>%
  unnest(glance) %>% 
  unnest(tidy, names_repair = "universal") %>% 
  dplyr::select(installation, index, statistic...8, r.squared...5, p.value...9, logLik...11) %>%
  rename(statistic = statistic...8,
         r_squared = r.squared...5,
         p_value = p.value...9,
         log_lik = logLik...11) %>%
  arrange(desc(r_squared), .by_group = FALSE) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")
```

**Discussion**  
Best fit linear models: the annual "temperature/heat" index describes much of the variability in heat stress illness rates.  

The installations with the best fit models are **Fort Campbell, KY**, **Fort Benning, GA**, and **Fort Jackson, SC**. These are all sites with high rates of HSIs. Fort Bragg, NC is an exception that has high HSI rates, but is not listed high among best fit linear models.  

Best fit indices (R^2^ above 40%):  

- days_wbgt_max_gt85pct (relative)
- days_wbgt_max_gt1sd (relative)  
- days_tmp_gt85pct (relative)  
- days_wbgt_gt82  
- hours_wbgt_gt82   
- days_wbgt_gt85  
- days_hi_gt80  
- days_wbgt_max_gt85pct (relative)  
- days_heat_index_gt85pct (relative)  
- days_heat_index_max_gt1sd (relative)  
- days_wbgt_max_gt90pct (relative)    
- days_wbgt_max_gt2sd_may_sep (relative)    
- days_heat_index_max_gt90pct (relative)    
- wbgt_f_max_anomaly_mean_may_sep (relative)    
- days_tmp_gt90pct_may_sep (relative)  
- days_heat_index_gt90pct_may_sep (relative)  
- days_wbgt_max_gt1sd_may_sep (relative)  
  
The top models are relative indices (based on 1990-2018 climatology), aggregated over the entire calendar year.  

WBGT models generally fit better than Heat Index or temperature.  



## Linear Models Summary, by Installation
```{r lm_models_grouped_sort}
# Model Summary Tables by Installation

annual_hsi_nest %>%
  unnest(glance) %>% 
  unnest(tidy, names_repair = "universal") %>% 
  dplyr::select(installation, index, statistic...8, r.squared...5, p.value...9, logLik...11) %>%
  rename(statistic = statistic...8,
         r_squared = r.squared...5,
         p_value = p.value...9,
         log_lik = logLik...11) %>%
  group_by(installation, add = TRUE) %>% 
  arrange(desc(r_squared), .by_group = TRUE) %>% 
  group_split() %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "300px")
```

Note: NA's are present for models where the conditions were not met (e.g. no days with HI above 125°F, no days with mean temp above 95th percentile). 

Installations with poorly fitting models (all R^2^ below 25%) are:  
**Fort Riley, KS** (highest R^2^ = 11.8%), **Fort Bragg, NC** (17.0%), **Fort Leonard Wood, MO** (19.0%), and **Fort Hood, TX** (22.8%).  

A WBGT-based model fit best for 6 installations, and a Heat Index-based model fit best for 5 (Hood, Jackson, Leonard Wood, Stewart).  


## Rank top performing models across all installations

```{r ranked_models}
# All models ranked by R^2
annual_hsi_nest %>%
  unnest(glance) %>% 
  unnest(tidy, names_repair = "universal") %>% 
  dplyr::select(installation, index, statistic...8, r.squared...5, p.value...9, logLik...11) %>%
  rename(statistic = statistic...8,
         r_squared = r.squared...5,
         p_value = p.value...9,
         log_lik = logLik...11) %>% 
  ungroup() %>% 
  arrange(desc(r_squared)) %>% 
    knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "300px")
  

# Select top models

annual_hsi_nest %>%
  unnest(glance) %>% 
  unnest(tidy, names_repair = "universal") %>% 
  dplyr::select(installation, index, statistic...8, r.squared...5, p.value...9, logLik...11) %>%
  rename(statistic = statistic...8,
         r_squared = r.squared...5,
         p_value = p.value...9,
         log_lik = logLik...11) %>% 
  ungroup() %>% 
  arrange(desc(r_squared)) %>% 
  dplyr::select(index) %>% 
  slice(1:12) %>% 
  distinct() %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 12)
  
```


## Selected indices (ambulatory)

Based on fits of linear models of annual heat illness ambulatory rates at 11 installations, we selected 10 indices out of 86 tested. 
Both absolute and relative indices are included. These models generally represent thresholds at the lower end of the heat intensity scales - rather than extreme heat days. 9 of the 10 are based on the full calendar year, rather than just the hot season. 

Further analysis: will hospitalizations be closer tied to more extreme heat indices?

**Absolute**
days_wbgt_gt82          
hours_wbgt_gt82         
days_wbgt_gt85          
days_hi_gt80 

**Relative**
days_tmp_gt85pct        
days_wbgt_max_gt85pct   
days_wbgt_max_gt1sd
days_heat_index_gt85pct 
days_wbgt_max_gt90pct   
days_tmp_gt90pct_may_sep


## Full / Pooled Model

Ambulatory rates, pooled: unweighted by base populations, no indicator variables for installation

```{r full_model, warning = FALSE}

joined_rate %>% 
  filter(index %in% c("days_hi_gt80", "days_wbgt_gt82", "hours_wbgt_gt82",
                      "days_wbgt_gt85", "days_tmp_gt85pct", "days_wbgt_max_gt85pct",
                      "days_wbgt_max_gt1sd", "days_heat_index_gt85pct", 
                      "days_wbgt_max_gt90pct", "days_tmp_gt90pct_may_sep")) %>% 
    ggplot(aes(x = value, y = rate)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.5, color = "black") +
    geom_smooth(se = FALSE, size = 0.5) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Ambulatory Rates and Annual Temperature/Heat Indices \n 
      Pooled across 11 installations (unweighted)",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 10)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

# Nest by index of heat

index_nest <-
  joined_rate %>%
    dplyr::select(index, everything()) %>% 
    nest(data = installation:rate) %>% 
    mutate(linear_models = map(data, annual_lm),
           glance = map(linear_models, broom::glance),
           tidy =   map(linear_models, broom::glance))
           
 
index_nest %>%
  unnest(glance) %>% 
  unnest(tidy, names_repair = "universal") %>% 
  dplyr::select(index, statistic...7, r.squared...4, p.value...8, logLik...10) %>%
  rename(statistic = statistic...7,
         r_squared = r.squared...4,
         p_value = p.value...8,
         log_lik = logLik...10) %>%
  arrange(desc(r_squared)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")


```

Poor fits of pooled, unweighted ambulatory annual rates of heat illness


## Pooled linear models, weighted by population

Hypothesis: poor fits due to weighting of Fort Bragg and Fort hood, which had poor fits in separate installation models.

```{r pooled_weighted}

# Join Army population by installation and year

dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Ambulatory Data",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
  dplyr::select(location, year, population) %>% 
  ggplot(aes(x = year, y = population, color = location)) +
    geom_point() +
    geom_line() + 
    labs(
      title = "Army Installation Population",
      x = "Year",
      y = "Population"
    ) +
  theme_bw()


dmed_population <-
  dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Ambulatory Data",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
  dplyr::select(location, year, population) %>% 
    mutate(location = recode(location, jackson = "fort_jackson"),
           location = recode(location, benning = "fort_benning_ga"),                 
           location = recode(location, bragg = "fort_bragg"),
           location = recode(location, campbell = "fort_campbell"),
           location = recode(location, polk = "fort_polk"),
           location = recode(location, hood = "fort_hood"),
           location = recode(location, stewart = "fort_stewart"),
           location = recode(location, leonardwood = "fort_leonard_wood"),
           location = recode(location, riley = "fort_riley"),
           location = recode(location, irwin = "ntc_and_fort_irwin"),
           location = recode(location, bliss = "fort_bliss"),
           location = as.character(location))





joined_rate %>%
  filter(index %in% c("days_hi_gt80", "days_wbgt_gt82", "hours_wbgt_gt82",
                      "days_wbgt_gt85", "days_tmp_gt85pct", "days_wbgt_max_gt85pct",
                      "days_wbgt_max_gt1sd", "days_heat_index_gt85pct", 
                      "days_wbgt_max_gt90pct", "days_tmp_gt90pct_may_sep")) %>%
  left_join(dmed_population, by = c("installation" = "location", "year")) %>% 
    ggplot(aes(x = value, y = rate, size = population)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = lm, se = TRUE, size = 0.5, color = "black", mapping = aes(weight = population)) +
    facet_wrap(~ index, scales = "free_x") +
    labs(
      title = "Army Heat Stress Illness Ambulatory Rates and Annual Temperature/Heat Indices \n 
      Pooled across 11 installations (weighted by populatio)",
      x = "Index value",
      y = "HSI rate (per 1,000 persons per year)"
    ) +
    theme_bw() +
    theme(strip.text = element_text(size = 10)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))


  

# Weighted linear regression function

weighted_lm = function(df) {
  df %>% 
    lm(rate ~ value, weights = population, data = .)
}




weighted_nest <-
joined_rate %>% 
  left_join(dmed_population, by = c("installation" = "location", "year")) %>% 
  dplyr::select(index, everything()) %>% 
    nest(data = installation:population) %>% 
    mutate(linear_models = map(data, weighted_lm),
           glance = map(linear_models, broom::glance),
           tidy =   map(linear_models, broom::glance))
  
weighted_nest %>%
  unnest(glance) %>% 
  unnest(tidy, names_repair = "universal") %>% 
  dplyr::select(index, statistic...7, r.squared...4, p.value...8, logLik...10) %>%
  rename(statistic = statistic...7,
         r_squared = r.squared...4,
         p_value = p.value...8,
         log_lik = logLik...10) %>%
  arrange(desc(r_squared)) %>% 
  knitr::kable(digits = 4) %>% 
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "700px")


```









**Discussion**  





