---
title: "installation_nldas"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(sf)
library(tidyverse)
library(viridis)
library(rvest)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(lwgeom)
library(ncdf4)
library(devtools)
library(raster)
library(ncdump)
library(lubridate)
library(flipTime)
library(fuzzyjoin)
library(RNetCDF)
library(tidync)
library(furrr)
library(tabularaster)
library(purrrlyr)
library(janitor)
library(zoo)
library(ggrepel)
library(broom)


```


# Installation NLDAS

This code links NLDAS estimates assembled in [`extract_cells`](analysis/extract_cells.html) to specific locations and summarizes the data.  

## Extract NLDAS from list for each location

```{r import_rds, echo = FALSE}
cell_key <- readRDS("data/cell_key.rds")
nldas_list <- read_rds("data/nldas_list.rds")
nldas_list_joint <- read_rds("data/nldas_list_joint.rds")
```

The NLDAS data are stored in two lists.  

The first set (`nldas_list`) contains:  
```{r first_set}

first_set_bases <- c(
      "Fort Benning GA", "Fort Bragg","Fort Campbell",
      "Fort Jackson", "Fort Polk", "Fort Hood",
      "Fort Stewart", "Fort Leonard Wood", "Fort Riley",
      "NTC and Fort Irwin", "Fort Bliss", "Fort Lewis",
      "Fort Sill", "Fort Huachuca", "Fort Carson", "Fort Gordon",
      "West Point Mil Reservation", "Fort Drum") %>% 
  stringr::str_replace_all(., " ", "_") %>%
  stringr::str_to_lower()

# West Point and Fort Huachica were initially extracted, but are not intended to be included in the "Top 25" analysis.

first_set_bases
```

The second set (`nldas_list_joint`) contains:
```{r second_set}

second_set_bases <- c(
           "MCB Quantico", "MCRD San Diego", "Twentynine Palms Main Base",
           "MCB Camp Lejeune", "MCRD Beaufort Parris Island", "MCB Camp Pendleton",
           "Eglin AFB", "Lackland AFB", "Fort Sam Houston") %>% 
  stringr::str_replace_all(., " ", "_") %>%
  stringr::str_to_lower()
  

second_set_bases 
```

Extract by location from first list and create an NLDAS object for each.


```{r first_extract, eval = FALSE}

for (i in seq_along(first_set_bases)) {

nldas_list %>% 
    lapply(function(x) x[i,]) %>% 
    bind_rows(., .id = "date_time") %>% 
  write_rds(., path = paste0("data/base_nldas/",  first_set_bases[i], ".rds" ))
}  

```

Extract by location from second list and create an NLDAS object for each.

```{r second_extract, eval = FALSE}

for (i in seq_along(second_set_bases)) {

nldas_list_joint %>% 
    lapply(function(x) x[i,]) %>% 
    bind_rows(., .id = "date_time") %>% 
  write_rds(., path = paste0("data/base_nldas/",  second_set_bases[i], ".rds" ))
}  

```

Review installation NLDAS files.
```{r check_saved_files}

list.files("data/base_nldas")

# Fort Benning (1st set)
read_rds("data/base_nldas/fort_benning_ga.rds")
#Eglin AFB (2nd set)
read_rds("data/base_nldas/eglin_afb.rds")

```


## Assemble master data frame with core variables

```{r combined_dataframe, warning = FALSE, eval = FALSE}
# Create index to substitute installation name into 
# master  dataframe id

df_index <- 
  list.files("data/base_nldas") %>%
      .[-c(10, 27)] %>% 
    as_tibble() %>% 
    mutate(row_id = row_number(),
           value = stringr::str_replace(value, ".rds", ""))

# Master dataframe (wide format)

nldas_df <-
  list.files("data/base_nldas", full.names = TRUE) %>%
    .[-c(10, 27)] %>%   # remove West Point and Fort Huachuca
    map_dfr(.x = ., .f = read_rds, .id = "id") %>%
  mutate(id = as.integer(id)) %>% 
  left_join(df_index, by = c("id" = "row_id")) %>% 
  rename(installation = value) %>% 
  dplyr::select(-id)

nldas_df

# write_rds(nldas_df, path = "data/nldas_df.rds")

```

```{r load_nldas_df}

nldas_df <- read_rds("data/nldas_df.rds")
nldas_df

```


