---
title: "land_cover"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(here)
library(sf)
library(sp)
library(tidyverse)
library(viridis)
library(rvest)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(rgeos)
library(lwgeom)
library(cowplot)
library(purrr)
library(purrrlyr)
library(tmap)
library(janitor)
library(plotly)
library(FedData)
library(tabularaster)
library(rasterVis)
library(raster)
```

## Load Military Installations, Ranges, and Training Areas (MIRTA) Dataset

Accessed from: https://catalog.data.gov/dataset/military-installations-ranges-and-training-areas.   
Metadata updated date: January 18, 2017.

Join Fort Benning Georgia and Alabama shapefiles. Note: Fort Lewis (JBLM) geometry includes remote sites.


## Select installations

Select bases for analyses. Citeria: in list of top CONUS bases for heat illnesses from MSMR report (April 2018). Additional Army installations based on HSI counts and rates (from DMED) and for diversity of climate regions.

Add centroid column.
```{r select_sf_dataframe, warning = FALSE}

# saveRDS(select_bases, "data/select_bases.rds")

select_bases <-
  read_rds("data/select_bases.rds")

army_shapefiles <-
  select_bases %>% 
  dplyr::filter(.$site_name %in% c("Fort Benning GA", "Fort Bragg","Fort Campbell",
                              "Fort Jackson", "Fort Polk", "Fort Hood",
                              "Fort Stewart", "Fort Leonard Wood", "Fort Riley",
                              "NTC and Fort Irwin", "Fort Bliss"))
```



## Download and crop the National Land Cover Database
ref: https://github.com/ropensci/FedData/blob/master/R/NLCD_FUNCTIONS.R

```{r get_nlcd, eval = FALSE}


canopy_list <- list()

for (i in seq_len(nrow(army_shapefiles))) {
    base_canopy <-
      get_nlcd(template = army_shapefiles$geometry[i],
                   year = 2016,
                   dataset = "Tree_Canopy",
                   label = army_shapefiles$site_name[i])  
    canopy_list[[i]] <- base_canopy
}



# write_rds(canopy_list, "data/canopy_list.rds")

```


## Tree Canopy Maps

```{r plot_tree_cover}

canopy_list <-
  read_rds("data/canopy_list.rds")

# Change canopy raster projection to base shapefile projection

projection <- "+proj=longlat +datum=WGS84 +no_defs"



# Function to canopy raster projections to base shapefile projections
canopy_crs_function <- function(stack) {
    projectRaster(stack, crs = projection)
}



canopy_list_crs <- 
  map(canopy_list, canopy_crs_function)
canopy_list_crs


plot_raster_function <- function(stack, base) {
    plot(stack, main = NULL)
    par(mar = c(2,2,2,2) + 0.1)
    plot(base, add = TRUE, lwd = 3)
}


map2(.x = canopy_list_crs, .y = army_shapefiles$geometry, .f = plot_raster_function)


```

## Mask raster by installation boundary


```{r mask_raster, eval = FALSE}

mask_canopy_list <- list()

for (i in seq_len(nrow(army_shapefiles))) {
    mask_canopy <-
        mask(canopy_list_crs[[i]], as(army_shapefiles$geometry[i], "Spatial"))
    
    mask_canopy_list[[i]] <- mask_canopy
}

names(mask_canopy_list) <- army_shapefiles$site_name
mask_canopy_list

purrr::map2(mask_canopy_list, names(mask_canopy_list), 
     ~plot(.x, main = .y))


# write_rds(mask_canopy_list, "data/mask_canopy_list.rds")

```



## Raster cell stats
sum, mean, min, max, sd
```{r raster_stats}

mask_canopy_list <-
  read_rds("data/mask_canopy_list.rds")


# Mean percent canopy at each installation
map_dfr(mask_canopy_list, ~ cellStats(., "mean")) %>% 
  tidyr::pivot_longer(., 1:11, names_to = "location",  values_to = "mean_pct_canopy")


# Canopy pixel frequency tables
# map(mask_canopy_list, freq)


# Canopy raster value statitics
mask_canopy_stats_list <- list()

for (i in seq_len(nrow(army_shapefiles))) {
    mask_canopy_stats <-
       freq(mask_canopy_list[[i]]) %>% 
       as_tibble() %>%
       summarise(sum_pixel = sum(count)) %>% 
       mutate(na_count = freq(mask_canopy_list[[i]]) %>%
                    as_tibble() %>% 
                    dplyr::filter(value %in% NA) %>% 
                    pull(count),
              pct_na = na_count / sum_pixel,
              mean_canopy = cellStats(mask_canopy_list[[i]], "mean"),
              sd_canopy = cellStats(mask_canopy_list[[i]], "sd"))
    
mask_canopy_stats_list[[i]] <- mask_canopy_stats 
}


mask_canopy_stats_list %>% 
  bind_rows() %>% 
  mutate(installation = army_shapefiles$site_name)


# Plot percent NA by mean_canopy value (includes space outside of mask)
mask_canopy_stats_list %>% 
  bind_rows() %>% 
  ggplot(aes(x = mean_canopy, y = pct_na)) +
    geom_point() +
    geom_smooth(method = lm)


# Histograms of canopy pixel values

purrr::map2(mask_canopy_list, names(mask_canopy_list), 
     ~histogram(.x, main = .y))

# Map "NA" values
# Inclusive of inside mask and bounding box
map2(mask_canopy_list, names(mask_canopy_list),  ~ plot(is.na(.), main = .y))



```


## Convert NA to 0 (zero), then mask shapefile

Reclassify function
```{r reclassify_na}

canopy_list_crs[[1]] %>% plot()

canopy_list_crs[[1]] %>% 
  is.na()

# Function to replace NA values
replace_na_function <- function(raster_layer) {
  raster::reclassify(raster_layer, cbind(NA, 0))
}


canopy_list_no_na <-
  map(.x = canopy_list_crs, .f = replace_na_function)

canopy_list_no_na

canopy_list_no_na[[1]] %>% plot()
```

Mask shapefile

```{r mask_no_na}

mask_canopy_no_na_list <- list()

for (i in seq_len(nrow(army_shapefiles))) {
    mask_canopy_no_na <-
        mask(canopy_list_no_na[[i]], as(army_shapefiles$geometry[i], "Spatial"))
    
    mask_canopy_no_na_list[[i]] <- mask_canopy_no_na
}

names(mask_canopy_no_na_list) <- army_shapefiles$site_name
mask_canopy_no_na_list

purrr::map2(mask_canopy_no_na_list, names(mask_canopy_no_na_list), 
     ~plot(.x, main = .y))


# write_rds(mask_canopy_list, "data/mask_canopy_list.rds")
```

## Canopy cell statistics with NA's reclassified


```{r cell_stats_no_na}
# Canopy raster value statitics
mask_canopy_stats_no_na_list <- list()

for (i in seq_len(nrow(army_shapefiles))) {
    mask_canopy_stats_no_na <-
       freq(mask_canopy_no_na_list[[i]]) %>% 
       as_tibble() %>%
       summarise(sum_pixel = sum(count)) %>% 
       mutate(na_count = freq(mask_canopy_no_na_list[[i]]) %>%
                    as_tibble() %>% 
                    dplyr::filter(value %in% NA) %>% 
                    pull(count),
              pct_na = na_count / sum_pixel,
              mean_canopy = cellStats(mask_canopy_no_na_list[[i]], "mean"),
              sd_canopy = cellStats(mask_canopy_no_na_list[[i]], "sd"))
    
mask_canopy_stats_no_na_list[[i]] <- mask_canopy_stats_no_na 
}


mask_canopy_stats_no_na_list %>% 
  bind_rows() %>% 
  mutate(installation = army_shapefiles$site_name)


# Histograms of canopy pixel values with NAs reclassified to 0

purrr::map2(mask_canopy_no_na_list, names(mask_canopy_list), 
     ~histogram(.x, main = .y))

```

