---
title: "mixed_models_poisson"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(readxl)
library(lubridate)
library(viridis)
library(knitr)
library(kableExtra)
library(zoo)
library(purrr)
library(furrr)
library(table1)
library(furniture)
library(visdat)
library(naniar)
library(modelr)
library(lme4)
library(ICC) 
library(dotwhisker)
library(broom)
library(car)
library(reshape2)
library(sjPlot)
library(ggridges)
library(readxl)

```

Annual models from DMED data.


Scope:

  CONUS Army Installations:  
  
    Fort Jackson, SC (Basic Combat Training site)     
    Fort Benning, GA (Basic Combat Training site)     
    Fort Bragg, NC  
    Fort Campbell, KY  
    Fort Polk, LA  
    Fort Hood, TX  
    Fort Stewart, GA  
    Fort Leonard Wood, MO (Basic Combat Training site)  
    Fort Riley, KS  
    Fort Irwin, CA  
    Fort Bliss, TX  
      
Population: Active-duty Army service members  

**Outcome**:
  Ambulatory (Out-patient), 1997 - 2018 / Hospitalization (In-Patient), 1990-2018 / Reportable Events, 1995 - 2018  
  Annual Count / Rate of Heat Stress Illness (any type, primary diagnosis)  

**Exposure indices**:  
*“Absolute” indices*  
  Annual mean (full-year): temperature, heat index, WBGT  
  Annual heat risk days / hours - Heat index above 80 / 90 / 103 / 125 °F - WBGT above 82 / 85 / 88 / 90 °F  
  
*“Relative” indices* (averaged over full-year and heat season months) 
Annual mean daily anomaly: temperature, heat index, WBGT  
Annual maximum daily anomaly: temperature, heat index, WBGT  
Days mean temperature index above daily climate normal percentile (averaged over all hours of day)  
  - temperature, heat index, WBGT above 85th / 90th / 95th percentiles  
Days maximum temperature index above daily climate normal maximum percentile  
  - temperature, heat index, WBGT above 85th / 90th / 95th percentiles  
Days mean temperature index above Standard Deviation(s) of mean daily temperature climate normal  
  - temperature, heat index, WBGT above 1 or 2 standard deviations of daily normal  
Days maximum temperature index above Standard Deviation(s) of max daily temperature climate normal  
  - temperature, heat index, WBGT above 1 or 2 standard deviations of maximum daily normal  

```{r read_data}

# Outcomes
dmed_tidy <- 
  read_rds("data/dmed_tidy.rds")

dmed_tidy

# Exposures
annual_tables_list <-
  read_rds("data/annual_tables_list.rds")

annual_tables_list

index_heat_key <-
  readxl::read_excel("data/index_heat_key.xlsx")

index_heat_key
# write_rds(index_heat_key, "data/index_heat_key.rds")


# Medical encounters - all ICD codes

all_icd <-
  read_rds("data/all_icd.rds")

```


```{r joined_rate_population}

dmed_population <-
  dmed_tidy %>% 
  filter(!location %in% c("us", "overseas"),
         type == "Ambulatory Data",
         hsi == "all",
         strata == "gender",
         category == "Total") %>% 
  dplyr::select(location, year, population) %>% 
    mutate(location = dplyr::recode(location, jackson = "fort_jackson"),
           location = dplyr::recode(location, benning = "fort_benning_ga"),                 
           location = dplyr::recode(location, bragg = "fort_bragg"),
           location = dplyr::recode(location, campbell = "fort_campbell"),
           location = dplyr::recode(location, polk = "fort_polk"),
           location = dplyr::recode(location, hood = "fort_hood"),
           location = dplyr::recode(location, stewart = "fort_stewart"),
           location = dplyr::recode(location, leonardwood = "fort_leonard_wood"),
           location = dplyr::recode(location, riley = "fort_riley"),
           location = dplyr::recode(location, irwin = "ntc_and_fort_irwin"),
           location = dplyr::recode(location, bliss = "fort_bliss"),
           location = as.character(location))


index_long <- 
  annual_tables_list %>% 
    filter(installation %in% c("fort_benning_ga", "fort_bragg", "fort_campbell", "fort_jackson",
                               "fort_polk", "fort_hood", "fort_stewart", "fort_leonard_wood", "fort_riley",
                               "ntc_and_fort_irwin", "fort_bliss")) %>% 
    unnest(data) %>% 
    pivot_longer(., cols = `1990`:`2019`, names_to = "year", values_to = "value") %>% 
    mutate(year = as.integer(year)) 


hsi_rates <-
  dmed_tidy %>% 
    filter(location %in% c("benning", "bragg", "campbell", "jackson",
                           "polk", "hood", "stewart", "leonardwood", 
                           "riley", "irwin", "bliss"),
           type == "Ambulatory Data",
           hsi == "all",
           strata == "gender",
           category == "Total") %>% 
    dplyr::select(location, year, rate) %>% 
    dplyr::mutate(location = dplyr::recode(location, jackson = "fort_jackson"),
           location = dplyr::recode(location, benning = "fort_benning_ga"),                 
           location = dplyr::recode(location, bragg = "fort_bragg"),
           location = dplyr::recode(location, campbell = "fort_campbell"),
           location = dplyr::recode(location, polk = "fort_polk"),
           location = dplyr::recode(location, hood = "fort_hood"),
           location = dplyr::recode(location, stewart = "fort_stewart"),
           location = dplyr::recode(location, leonardwood = "fort_leonard_wood"),
           location = dplyr::recode(location, riley = "fort_riley"),
           location = dplyr::recode(location, irwin = "ntc_and_fort_irwin"),
           location = dplyr::recode(location, bliss = "fort_bliss"),
           location = as.character(location))

# Ambulatory HSI

hsi_amb <-
  dmed_tidy %>% 
    filter(location %in% c("benning", "bragg", "campbell", "jackson",
                           "polk", "hood", "stewart", "leonardwood", 
                           "riley", "irwin", "bliss"),
           type == "Ambulatory Data",
           hsi == "all",
           strata == "gender",
           category == "Total") %>% 
    dplyr::select(location, year, count, population) %>% 
    dplyr::mutate(location = dplyr::recode(location, jackson = "fort_jackson"),
           location = dplyr::recode(location, benning = "fort_benning_ga"),                 
           location = dplyr::recode(location, bragg = "fort_bragg"),
           location = dplyr::recode(location, campbell = "fort_campbell"),
           location = dplyr::recode(location, polk = "fort_polk"),
           location = dplyr::recode(location, hood = "fort_hood"),
           location = dplyr::recode(location, stewart = "fort_stewart"),
           location = dplyr::recode(location, leonardwood = "fort_leonard_wood"),
           location = dplyr::recode(location, riley = "fort_riley"),
           location = dplyr::recode(location, irwin = "ntc_and_fort_irwin"),
           location = dplyr::recode(location, bliss = "fort_bliss"),
           location = as.character(location),
           rate = (count / population) * 1000)


# Hospitalization HSI

hsi_hosp <-
  dmed_tidy %>% 
    filter(location %in% c("benning", "bragg", "campbell", "jackson",
                           "polk", "hood", "stewart", "leonardwood", 
                           "riley", "irwin", "bliss"),
           type == "Hospitalizations",
           hsi == "all",
           strata == "gender",
           category == "Total") %>% 
    dplyr::select(location, year, count, population) %>% 
    dplyr::mutate(location = dplyr::recode(location, jackson = "fort_jackson"),
           location = dplyr::recode(location, benning = "fort_benning_ga"),                 
           location = dplyr::recode(location, bragg = "fort_bragg"),
           location = dplyr::recode(location, campbell = "fort_campbell"),
           location = dplyr::recode(location, polk = "fort_polk"),
           location = dplyr::recode(location, hood = "fort_hood"),
           location = dplyr::recode(location, stewart = "fort_stewart"),
           location = dplyr::recode(location, leonardwood = "fort_leonard_wood"),
           location = dplyr::recode(location, riley = "fort_riley"),
           location = dplyr::recode(location, irwin = "ntc_and_fort_irwin"),
           location = dplyr::recode(location, bliss = "fort_bliss"),
           location = as.character(location),
           rate = (count / population) * 1000) 



# Reportable Events Counts

hsi_rme <-
  dmed_tidy %>% 
    filter(location %in% c("benning", "bragg", "campbell", "jackson",
                           "polk", "hood", "stewart", "leonardwood", 
                           "riley", "irwin", "bliss"),
           type == "Reportable Events",
           hsi == "all",
           strata == "gender",
           category == "Total") %>% 
    dplyr::select(location, year, count, population) %>% 
    dplyr::mutate(location = dplyr::recode(location, jackson = "fort_jackson"),
           location = dplyr::recode(location, benning = "fort_benning_ga"),                 
           location = dplyr::recode(location, bragg = "fort_bragg"),
           location = dplyr::recode(location, campbell = "fort_campbell"),
           location = dplyr::recode(location, polk = "fort_polk"),
           location = dplyr::recode(location, hood = "fort_hood"),
           location = dplyr::recode(location, stewart = "fort_stewart"),
           location = dplyr::recode(location, leonardwood = "fort_leonard_wood"),
           location = dplyr::recode(location, riley = "fort_riley"),
           location = dplyr::recode(location, irwin = "ntc_and_fort_irwin"),
           location = dplyr::recode(location, bliss = "fort_bliss"),
           location = as.character(location),
           rate = (count / population) * 1000)



# Join with Annual heat; assign Basic Combat Training (BCT) category

# Ambultatory

joined_hsi_amb <-
  index_long %>% 
    left_join(hsi_amb, by = c("installation" = "location", "year" = "year")) %>% 
      drop_na(count) %>% 
  mutate(bct = case_when(installation %in% c("fort_benning_ga", "fort_jackson", "fort_leonard_wood") ~ 1,
                         !installation %in% c("fort_benning_ga", "fort_jackson", "fort_leonard_wood") ~ 0))  %>%
  ungroup() %>% 
  dplyr::group_by(index) %>% 
      mutate(value_scaled = scale(value, center = TRUE, scale = TRUE)) %>% 
  ungroup() %>% 
  na.omit()



# Hospitalization
  # Join to indices of heat; assign BCT status; center and scale heat index value (pooled, for each index)  


joined_hsi_hosp <-
  index_long %>% 
    left_join(hsi_hosp, by = c("installation" = "location", "year" = "year")) %>% 
      drop_na(count) %>% 
  mutate(bct = case_when(installation %in% c("fort_benning_ga", "fort_jackson", "fort_leonard_wood") ~ 1,
                         !installation %in% c("fort_benning_ga", "fort_jackson", "fort_leonard_wood") ~ 0)) %>%
  ungroup() %>% 
  dplyr::group_by(index) %>% 
      mutate(value_scaled = scale(value, center = TRUE, scale = TRUE)) %>% 
  ungroup() %>% 
  na.omit()

# write_rds(joined_hsi_hosp, "data/joined_hsi_hosp.rds", compress = "none")



# Reportable Events
joined_hsi_rme <-
  index_long %>% 
    left_join(hsi_rme, by = c("installation" = "location", "year" = "year")) %>% 
      drop_na(count) %>% 
  mutate(bct = case_when(installation %in% c("fort_benning_ga", "fort_jackson", "fort_leonard_wood") ~ 1,
                         !installation %in% c("fort_benning_ga", "fort_jackson", "fort_leonard_wood") ~ 0)) %>%
  ungroup() %>% 
  dplyr::group_by(index) %>% 
      mutate(value_scaled = scale(value, center = TRUE, scale = TRUE)) %>% 
  ungroup() %>% 
  na.omit()

# write_rds(joined_hsi_rme, "data/joined_hsi_rme.rds", compress = "none")



```


## Plots of HSI Ambulatory Counts

```{r plot_ambulatory_rates, cache = TRUE, warning = FALSE}

joined_hsi_amb %>% 
 ggplot(aes(x = year, y = population, color = installation, shape = installation)) +
    geom_line() +
    geom_point() +
    labs(
      title = "Population",
      y = "Population",
      x = "Year"
    ) +
    scale_shape_manual(values = 0:11) +
    theme_bw() 


joined_hsi_amb %>% 
 ggplot(aes(x = count)) +
    geom_histogram(color = "darkblue", fill = "lightblue") +
    facet_wrap(installation ~ .) +
    labs(
      title = "Army Heat Stress Illness Ambulatory Counts, 1997-2018",
      y = "Installation",
      x = "HSI count"
    ) +
    theme_bw() 


```

## Examine single index of heat (out of 80+ indices)  

Annual hours WBGT greater than 85 deg F
```{r}


joined_hsi_amb %>% 
  filter(index %in% "hours_wbgt_gt85") %>% 
  ggplot(aes(x = year, y = count)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.8) +
    labs(
      title = "Army Heat Stress Illness Ambulatory Counts",
      x = "Year",
      y = "HSI count"
    ) +
    theme_bw() 


joined_hsi_amb %>% 
  filter(index %in% "hours_wbgt_gt85") %>% 
  ggplot(aes(x = year, y = count, color = installation, shape = installation)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.8) +
    labs(
      title = "Army Heat Stress Illness Ambulatory Counts",
      x = "Year",
      y = "HSI count"
    ) +
    theme_bw() 


joined_hsi_amb %>% 
  filter(index %in% "hours_wbgt_gt85") %>% 
  ggplot(aes(x = value, y = count, color = installation, shape = installation)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE, size = 0.8) +
    labs(
      title = "Army Heat Stress Illness Ambulatory Counts at 11 Installations (1997-2018)",
      x = "Hours WBGT > 85 deg F",
      y = "HSI count"
    ) +
    scale_shape_manual(values = 0:11) +
    theme_bw() 

joined_hsi_amb %>% 
   filter(index %in% "hours_wbgt_gt85") %>% 
 ggplot(aes(x = value, y = fct_reorder(installation, value))) +
    ggridges::geom_density_ridges(
       jittered_points = TRUE,
       aes(point_color = installation, point_fill = installation), 
       alpha = 0.5,
       point_alpha = 1) +
    labs(
      title = "Hours WBGT exceeding 85 deg F, 1997-2018",
      y = "Installation",
      x = "Hours WBGT > 85"
    ) +
    theme_bw() +
    theme(legend.position = "none")
```

## Observational Findings

1. Ambulatory HSI rates increased over time.
2. WBGT temperatures increased over time.
3. Slopes of HSI rate change vary by installation.
4. Slopes of WBGT increase are consistent by installation (clusters below 100 hours/year and above 200 hours/year).
5. HSI rates increase with the number of hours of WBGT exceeding 85 deg F.
6. Slopes of HSI rate increases with hours WBGT > 85 vary by installation (among installations with the highest HSI rates). 

## Models

```{r scaled}

# Assign dataset; center and scale predictor value

data_hours_wbgt_gt85_counts <-
  joined_hsi_amb %>% 
    filter(index %in% "hours_wbgt_gt85")


```

```{r models_initial, eval = FALSE}
# Fitting ‘installation’ as a fixed effect in model M1 assumes 
# the 11 ‘installation’ means are all independent of one another, 
# and share a common residual variance

pooled <- glm(count ~ value + installation, data = data_hours_wbgt_gt85_counts)
summary(pooled)

pooled_offset <- glm(count ~ value + installation, data = data_hours_wbgt_gt85_counts, offset = log(population))
summary(pooled_offset)


# Fitting installation as a random intercept model in model M2 assumes that the 1 measured
# installation means are only a subset of the realised possibilities drawn from a ‘global’ set of
# population means that follow a Normal distribution with its own mean and variance.

random_intercept <- glmer(count ~ value_scaled + (1|installation), data = data_hours_wbgt_gt85_counts, family = poisson, offset = log(population))

# Fitting random intercepts and slopes allows the slope of a 
# predictor to vary based on a separate grouping variable.

poisson_mod <- glm(count ~ value_scaled + installation, data = data_hours_wbgt_gt85_counts, family = poisson)
sjPlot::tab_model(poisson_mod, p.val = "kr", show.df = TRUE)

poisson_mod_offset <- glm(count ~ value_scaled + installation, data = data_hours_wbgt_gt85_counts, family = poisson, offset = log(population))
sjPlot::tab_model(poisson_mod, poisson_mod_offset, p.val = "kr", show.df = TRUE)

random_intercept_slope_no_offset <- glmer(count ~ value_scaled + (1 + value_scaled|installation), data = data_hours_wbgt_gt85_counts, family = poisson)
summary(random_intercept_slope_no_offset)


```

## Poisson with population offset, mixed effects (random intercepts and slopes)

```{r models_selected}

random_intercept_slope <- glmer(count ~ value_scaled + (1 + value_scaled|installation), data = data_hours_wbgt_gt85_counts, family = poisson, offset = log(population))

summary(random_intercept_slope)
sjPlot::tab_model(random_intercept_slope, p.val = "kr", show.df = TRUE)


random_intercept_slope_bct <- glmer(count ~ value_scaled + bct + (1 + value_scaled|installation), data = data_hours_wbgt_gt85_counts, family = poisson, offset = log(population))

summary(random_intercept_slope_bct)
broom::glance(random_intercept_slope_bct)
broom::tidy(random_intercept_slope_bct)
sjPlot::tab_model(random_intercept_slope_bct, p.val = "kr", show.df = TRUE)
```

## Plot models

```{r model_plots}

sjPlot::plot_model(random_intercept_slope_bct)
sjPlot::plot_model(random_intercept_slope_bct, type = "pred", terms = "value_scaled")
sjPlot::plot_model(random_intercept_slope_bct, type = "re", terms = "value_scaled")
sjPlot::plot_model(random_intercept_slope_bct, type = "resid", terms = "value_scaled")
sjPlot::plot_model(random_intercept_slope_bct, type = "diag", terms = "value_scaled")

```

## Table of index means and standard deviations (pooled)
For interpreting centered and scaled estimates.

```{r index_mean_sd}
index_mean_sd <- 
  joined_hsi_amb %>% 
    group_by(index) %>% 
    summarise(
      mean = mean(value),
      sd = sd(value)
    )

index_mean_sd

```


## Multiple Models Function

```{r glmer_function, warning = FALSE}


random_slope_bct_function = function(df) {
  df %>% 
    glmer(count ~ value_scaled + bct + (1 + value_scaled|installation), data = ., family = poisson, offset = log(population))
}

# Ambulatory nested models

random_slope_bct_nest <-
  joined_hsi_amb %>%
    dplyr::select(index, everything()) %>% 
      nest(data = installation:value_scaled) %>% 
      mutate(glmer_models = map(data, random_slope_bct_function),
             glance = map(glmer_models, broom::glance),
             tidy =   map(glmer_models, broom::tidy),
             augment = map(glmer_models, broom::augment))


random_slope_bct_nest 
random_slope_bct_nest$index %>% View()
# Check singular fit models

map(random_slope_bct_nest$glmer_models, isSingular) %>% 
  unlist() %>% 
  tibble::enframe(name = NULL) %>% 
  mutate(index = random_slope_bct_nest$index) %>% 
  filter(value %in% "TRUE")

# Remove `days_wbgt_gt95pct_may_sep`: singular fit model

random_slope_bct_nest <-
  random_slope_bct_nest %>% 
  filter(!index %in% "days_wbgt_gt95pct_may_sep")


# write_rds(random_slope_bct_nest, "data/random_slope_bct_nest.rds", compress = "none")



# Model summaries

# Arranged by magnitude of index value estimate

random_slope_bct_nest  %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>%
  filter(term %in% "value_scaled") %>% 
  arrange(desc(estimate))

# Arranged by magnitude of "bct" estimate

random_slope_bct_nest  %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>%
  filter(term %in% "bct") %>% 
  arrange(desc(estimate))


# Arranged by AIC (lower is better)

random_slope_bct_nest  %>% 
  unnest(glance, .drop = TRUE) %>% 
  dplyr::select(index, sigma:df.residual) %>% 
  arrange(AIC)



```



```{r}
# Ambulatory estimates plot

random_slope_bct_nest  %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>%
  filter(term %in% "value_scaled") %>%
  mutate(index = fct_reorder(index, estimate)) %>%
  ggplot() +
    geom_point(aes(x = estimate, y = index)) +
    geom_errorbarh(
      aes(y = index,
          xmin = estimate - 1.96 * std.error,
          xmax = estimate + 1.96 * std.error)) +
     geom_vline(aes(xintercept = 0), colour = "blue") +
     theme_bw() +
     theme(axis.text.y = element_text(size = rel(0.75))) +
      ggtitle("Regression coefficients - Ambulatory, value-scaled")
   

      
 random_slope_bct_nest  %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>%
  filter(term %in% "value_scaled") %>%
  mutate(index = fct_reorder(index, estimate)) %>% 
  ggplot() +
    geom_point(aes(x = exp(estimate), y = index)) +
    geom_errorbarh(
      aes(y = index,
          xmin = exp(estimate - 1.96 * std.error),
          xmax = exp(estimate + 1.96 * std.error))) +
     geom_vline(aes(xintercept = 1), colour = "blue") +
     theme_bw() +
     theme(axis.text.y = element_text(size = rel(0.75))) +
     ggtitle("Rate Ratio - Ambulatory, value-scaled")



# p-value boxplot
random_slope_bct_nest  %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>%
  filter(term %in% "value_scaled") %>% 
  ggplot(aes(x = "", y = p.value)) +
        geom_boxplot() +
        geom_hline(aes(yintercept = 0.05), colour = "blue") +
        geom_rug() +
        coord_flip() +
        ggtitle("Boxplot of value-scaled p-values") +
        theme_bw()

```





## Plot estimates of all nested models

```{r plot_all_models}

#random_slope_bct_nest %>% 
#  pull(glmer_models) %>% 
#  map(sjPlot::plot_model)
#  
#

# Plot 10

models_10_names <-
  random_slope_bct_nest  %>% 
      unnest(tidy, .drop = TRUE) %>% 
      dplyr::select(index, term:group) %>%
      filter(term %in% "value_scaled") %>% 
      arrange(desc(estimate)) %>% 
      slice(1:10) %>%
      pull(index)
           
models_10 <-
  random_slope_bct_nest %>% 
  filter(index %in% models_10_names) %>% 
  pull(glmer_models)

map2(.x = models_10, .y = models_10_names, .f = ~ sjPlot::plot_model(.x, title = .y))  
map2(.x = models_10, .y = models_10_names, .f = ~ sjPlot::plot_model(.x, type = "pred", terms = "value_scaled", title = .y))
map2(.x = models_10, .y = models_10_names, .f = ~ sjPlot::plot_model(.x, type = "re", terms = "value_scaled", title = .y))


  
index_mean_sd %>% 
  filter(index %in% "models_10_names") %>% 
  knitr::kable(digits = 3) %>% 
  kableExtra::kable_styling()



 random_slope_bct_nest %>% 
  filter(index %in% "hours_wbgt_gt82") %>% 
  pull(glmer_models) %>%  
  sjPlot::tab_model(p.val = "kr", show.df = TRUE)

 
 random_slope_bct_nest %>% 
  filter(index %in% "hours_wbgt_gt82") %>% 
  pull(glmer_models) %>%   
summary()
 
 
broom::glance(random_intercept_slope_bct)
broom::tidy(random_intercept_slope_bct)

  

  
```

## Hospitalization Nested
```{r}

# Hospitalization nested models

hosp_random_slope_bct_nest <-
  joined_hsi_hosp %>%
  na.omit() %>% 
    dplyr::select(index, everything()) %>% 
      nest(data = installation:value_scaled) %>% 
      mutate(glmer_models = map(data, random_slope_bct_function),
             glance = map(glmer_models, broom::glance),
             tidy =   map(glmer_models, broom::tidy),
             augment = map(glmer_models, broom::augment))

# Check singular fit models

map(hosp_random_slope_bct_nest$glmer_models, isSingular) %>% 
  unlist() %>% 
  tibble::enframe(name = NULL) %>% 
  mutate(index = hosp_random_slope_bct_nest$index) %>% 
  filter(value %in% "TRUE")

# Remove `days_tmp_gt85pct_may_sep`, `days_tmp_gt95pct_may_sep` , `days_wbgt_gt95pct_may_sep`: singular fit model

hosp_random_slope_bct_nest <-
  hosp_random_slope_bct_nest %>% 
  filter(!index %in% c("days_wbgt_gt95pct_may_sep", "days_tmp_gt85pct_may_sep", "days_tmp_gt95pct_may_sep"))

#write_rds(hosp_random_slope_bct_nest, "data/hosp_random_slope_bct_nest.rds", compress = "none")
```



```{r}
hosp_random_slope_bct_nest  %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>%
  filter(term %in% "value_scaled") %>%
  mutate(index = fct_reorder(index, estimate)) %>%
  ggplot() +
    geom_point(aes(x = estimate, y = index)) +
    geom_errorbarh(
      aes(y = index,
          xmin = estimate - 1.96 * std.error,
          xmax = estimate + 1.96 * std.error)) +
     geom_vline(aes(xintercept = 0), colour = "blue") +
     theme_bw() +
     theme(axis.text.y = element_text(size = rel(0.75))) +
      ggtitle("Regression coefficients - Hospitalization, value-scaled")
   

      
 hosp_random_slope_bct_nest  %>% 
  unnest(tidy, .drop = TRUE) %>% 
  dplyr::select(index, term:group) %>%
  filter(term %in% "value_scaled") %>%
  mutate(index = fct_reorder(index, estimate)) %>% 
  ggplot() +
    geom_point(aes(x = exp(estimate), y = index)) +
    geom_errorbarh(
      aes(y = index,
          xmin = exp(estimate - 1.96 * std.error),
          xmax = exp(estimate + 1.96 * std.error))) +
     geom_vline(aes(xintercept = 1), colour = "blue") +
     theme_bw() +
     theme(axis.text.y = element_text(size = rel(0.75))) +
     ggtitle("Rate Ratio - Hospitalization, value-scaled")


 
```

