---
title: "nldas_averaged"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(tidyverse)
library(dygraphs)
library(xts)
library(timetk)
library(zoo)
library(forecast)
library(fable)
library(sweep)
library(viridis)
library(tsibble)
library(lubridate)
```

We want to use local time for daily averages. 

We will modify duplicate times due to the end of daylight savings by adding one second to the repeated time.

```{r}
nldas_wide <- read_rds("data/nldas_wide.rds")

# Add one second to duplicate times
local_dst_adjust <-
  nldas_wide %>%
    mutate(row = dplyr::row_number()) %>% 
    tsibble::duplicates(key = installation, index = local_dttm) %>% 
    filter(local_dttm != lead(local_dttm, default = "1")) %>% 
    mutate(local_dttm = local_dttm + seconds(1))

local_dst_adjust

# Remove rows with duplicate local times and replace with updated rows (+1 second)

nldas_wide <-
  nldas_wide %>%
    mutate(row = dplyr::row_number()) %>% 
    filter(!row %in% local_dst_adjust$row) %>% 
    bind_rows(local_dst_adjust) %>% 
    arrange(local_dttm) %>% 
    dplyr::select(-row)

# Check for duplicates
nldas_wide %>% 
  duplicates(key = installation, index = local_dttm)


nldas_tsibble <-
  nldas_wide %>% 
    dplyr::select(installation, local_dttm, tmp_f, heat_index, wbgt_f) %>% 
    tsibble::as_tsibble(., key = installation, index = local_dttm)

```

## Aggregate over calendar periods  

### Averaged from hourly to daily
```{r nldas_daily}


nldas_daily <-
  nldas_tsibble %>% 
    tsibble::group_by_key() %>%
    tsibble::index_by(date = ~ as.Date(.)) %>% # daily aggregates
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(mean = "mean", max = "max", min = "min", sd = "sd")) %>% 
  filter(date >= date("1990-01-01"))



nldas_daily

nldas_daily %>% 
  filter(date >= date("2019-07-01")) %>% 
    autoplot(wbgt_f_mean)

```

### Averaged from hourly to monthly

```{r nldas_monthly}

nldas_monthly <-
  nldas_tsibble %>% 
    tsibble::group_by_key() %>%
    tsibble::index_by(year_month = ~ yearmonth(.)) %>% # daily aggregates
    dplyr::summarise_at(
          .vars = vars(tmp_f, heat_index, wbgt_f),
          .funs = c(mean = "mean", max = "max", min = "min", sd = "sd")) %>% 
  filter(year_month >= date("1990-01-01"))
  
nldas_monthly

nldas_monthly %>% 
  filter(year_month >= date("2018-01-01")) %>% 
    autoplot(wbgt_f_mean)

```

### Averaged from hourly to annual

```{r}
nldas_annual <-
  nldas_tsibble %>% 
    tsibble::group_by_key() %>%
    tsibble::index_by(year = ~ lubridate::year(.)) %>% # daily aggregates
    dplyr::summarise_at(
              .vars = vars(tmp_f, heat_index, wbgt_f),
              .funs = c(mean = "mean", max = "max", min = "min", sd = "sd")) %>% 
  filter(!year %in% "1989")
  
nldas_annual

# write_rds(nldas_annual, "data/nldas_annual.rds")

nldas_annual %>% 
    autoplot(wbgt_f_mean)
```

## Annual Heat Risk Days

Counts of days per year above Heat Index and WBGT category limits.

Heat Index groups: maximum daily heat index above 80, 90, 103, 125°F  

WBGT groups: maximum daily WBGT above 82, 85, 88, 90°F  

```{r, heat_risk_days}

heat_risk_days <-
  nldas_daily %>% 
    group_by(installation) %>% 
    index_by(year = lubridate::year(date)) %>% 
      summarise(
        days_hi_gt80 = sum(heat_index_max > 80),
        days_hi_gt90 = sum(heat_index_max > 90),
        days_hi_gt103 = sum(heat_index_max > 103),
        days_hi_gt125 = sum(heat_index_max > 125),
        days_wbgt_gt82 = sum(wbgt_f_max > 82),
        days_wbgt_gt85 = sum(wbgt_f_max > 85),
        days_wbgt_gt88 = sum(wbgt_f_max > 88),
        days_wbgt_gt90 = sum(wbgt_f_max > 90)
      )
  
heat_risk_days

# write_rds(heat_risk_days, "data/heat_risk_days.rds")

```

### Boxplots of heat risk day counts per year 
```{r heat_risk_boxplot}

heat_risk_days %>% 
  gather(key = "metric", value = "count", days_hi_gt80:days_wbgt_gt90) %>%
  filter(metric == c("days_hi_gt90", "days_wbgt_gt85")) %>%
  ggplot(aes(x = fct_reorder(installation, count), y = count, fill = metric)) +
    geom_boxplot() +
    coord_flip() +
    theme_bw() +
    theme(legend.position = "bottom")

```


### Bar chart heat risk day counts per year 
```{r heat_risk_plot_by_year, , fig.width = 12, fig.height = 24}

heat_risk_days %>%
  filter(!year %in% "1989") %>% 
  gather(key = "metric", value = "count", days_hi_gt80:days_wbgt_gt90) %>% 
  ggplot(aes(x = year, y = count, fill = metric)) +
    geom_col(position = "dodge") +
    facet_wrap(~ installation, ncol = 1) +
    theme_bw() +
    theme(strip.text.x = element_text(margin = margin(.1, 0, .1, 0, "cm"))) +
    theme(legend.position = "bottom")


```

### Annual heat risk days trend: Heat Index
```{r heat_risk_day_trend}

heat_risk_days %>%
  filter(!year %in% "1989") %>%
  gather(key = "metric", value = "count", days_hi_gt80:days_wbgt_gt90) %>%
  filter(metric == "days_hi_gt90") %>% 
  ggplot(aes(x = year, y = count, color = installation)) +
  geom_point() +
  geom_smooth(size = 0.5, method = lm, se = FALSE) +
  labs(
    title = "Heat Risk Days: Heat Index Above 90°F",
    x = "Year",
    y = "Days with Heat Index > 90°F") +
  theme_bw() +
  theme(legend.position = "bottom")
```

### Annual heat risk days trend: WBGT

```{r heat_risk_day_trend_wbgt}

heat_risk_days %>%
  filter(!year %in% "1989") %>%
  gather(key = "metric", value = "count", days_hi_gt80:days_wbgt_gt90) %>% 
  filter(metric == "days_wbgt_gt85") %>% 
  ggplot(aes(x = year, y = count, color = installation)) +
  geom_point() +
  geom_smooth(size = 0.5, method = lm, se = FALSE) +
  labs(
    title = "Heat Risk Days: WBGT Above 85°F",
    x = "Year",
    y = "Days with WBGT > 85°F") +
  theme_bw() +
  theme(legend.position = "bottom")

```


## Annual Heat Risk Hours

Counts of days per year above Heat Index and WBGT category limits.

Heat Index groups: hourly heat index above 80, 90, 103, 125°F  

WBGT groups: hourly WBGT above 82, 85, 88, 90°F  



```{r heat_risk_hours}

heat_risk_hours <-
  nldas_tsibble %>% 
    group_by(installation) %>% 
    index_by(year = lubridate::year(local_dttm)) %>% 
      summarise(
        hours_hi_gt80 = sum(heat_index > 80),
        hours_hi_gt90 = sum(heat_index > 90),
        hours_hi_gt103 = sum(heat_index > 103),
        hours_hi_gt125 = sum(heat_index > 125),
        hours_wbgt_gt82 = sum(wbgt_f > 82),
        hours_wbgt_gt85 = sum(wbgt_f > 85),
        hours_wbgt_gt88 = sum(wbgt_f > 88),
        hours_wbgt_gt90 = sum(wbgt_f > 90)
      )

heat_risk_hours

# write_rds(heat_risk_hours, "data/heat_risk_hours.rds")
```


### Boxplots of heat risk hour counts per year 
```{r heat_risk_hours_boxplot}

heat_risk_hours %>%
  filter(!year %in% "1989") %>% 
  gather(key = "metric", value = "count", hours_hi_gt80:hours_wbgt_gt90) %>%
  filter(metric == c("hours_hi_g90", "hours_wbgt_gt85")) %>%
  ggplot(aes(x = fct_reorder(installation, count), y = count, fill = metric)) +
    geom_boxplot() +
    coord_flip() +
    theme_bw() +
    theme(legend.position = "bottom")

```




