---
title: "nldas_indices_end2019"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(sf)
library(tidyverse)
library(lutz)
library(lubridate)
library(weathermetrics)
library(humidity)
library(wbgt)
library(HeatStress)
library(inspectdf)
library(janitor)

```


Expand dataframe with indices and unit conversions


## Identify timezones for local time conversion

```{r id_timezones, eval = FALSE}
# GMT adjust

time_zones <-
  read_rds("data/select_bases.rds") %>% 
  dplyr::select(site_name, centroid) %>% 
  mutate(time_zone = lutz::tz_lookup(centroid, crs = 4326, method = "accurate")) %>% 
  as_tibble() %>% 
  dplyr::select(site_name, time_zone) %>% 
  mutate(site_name = stringr::str_replace_all(site_name, " ", "_"),
         site_name =  stringr::str_to_lower(site_name))

time_zones  

lutz::tz_plot("America/New_York")
lutz::tz_plot("America/Chicago")
lutz::tz_plot("America/Denver")
lutz::tz_plot("America/Los_Angeles")


```

## Add local time column to NLDAS dataframe

This will serve as an input to WBGT estimate calculations.

```{r local_time, eval = FALSE}

nldas_df <- read_rds("data/nldas_df.rds")

# Inspect for missing values
nldas_df_end2019 %>% inspectdf::inspect_na()

# timezones cannot be vectorized in lubridate, therefore `case_when` used
# https://community.rstudio.com/t/working-with-timezones-in-lubridate/4260



# https://stackoverflow.com/questions/55352166/how-to-use-lubridate-as-datetime-function-in-combination-with-dplyr-mutate-and-c
# (Without the additional `as_datetime(as.character))`, each value receives the result of the first item in case_when- Eglin converts to EST instead of CST)


nldas_time_zone <-
  nldas_df_end2019 %>% 
    mutate(utc_dttm = lubridate::ymd_hm(date_time)) %>% 
    dplyr::select(utc_dttm, installation, everything(), -date_time) %>% 
    left_join(time_zones, by = c("installation" = "site_name")) %>%  
    dplyr::mutate(local_dttm = dplyr::case_when(
      .$time_zone == "America/New_York" ~ as.character(with_tz(.$utc_dttm, tz = "America/New_York")),
      .$time_zone == "America/Chicago" ~ as.character(with_tz(.$utc_dttm, tz = "America/Chicago")), 
      .$time_zone == "America/Denver" ~ as.character(with_tz(.$utc_dttm, tz = "America/Denver")), 
      .$time_zone == "America/Los_Angeles" ~ as.character(with_tz(.$utc_dttm, tz = "America/Los_Angeles"))),
      local_dttm = lubridate::ymd_hms(local_dttm)) %>% 
  dplyr::select(utc_dttm, local_dttm, everything(), -time_zone)
    

```


## Meteorology conversions and heat index
From original NLDAS variables, add windspeed, temp in F and C, relative humidity, dew point (C), heat index (F) 

```{r nldas_conversions, eval = FALSE}

nldas_wide_end2019 <-
  nldas_time_zone %>% 
    mutate(windspeed = sqrt(UGRD^2 + VGRD^2),
           tmp_f = weathermetrics::convert_temperature(TMP, old_metric = "kelvin", new_metric = "fahrenheit"),
           tmp_c = weathermetrics::convert_temperature(TMP, old_metric = "kelvin", new_metric = "celsius"),
           rh = humidity::SH2RH(SPFH, TMP, p = PRES, isK = TRUE),
           dp_c = humidity.to.dewpoint(rh = rh, t = tmp_c, temperature.metric = "celsius"),
           heat_index = weathermetrics::heat.index(t = tmp_f,
                           rh = rh,
                           temperature.metric = "fahrenheit",
                           output.metric = "fahrenheit",
                           round = 2)) 


# Inspect missing data: dew point/ heat index if relative humidity below 0% or above 100%

nldas_wide_end2019 %>% 
  inspectdf::inspect_na()

nldas_wide %>% 
  filter(is.na(dp_c)) %>% 
  arrange(desc(rh))

# Restrict RH to 100%

nldas_wide_end2019 <-
  nldas_time_zone %>% 
    mutate(windspeed = sqrt(UGRD^2 + VGRD^2),
           tmp_f = weathermetrics::convert_temperature(TMP, old_metric = "kelvin", new_metric = "fahrenheit"),
           tmp_c = weathermetrics::convert_temperature(TMP, old_metric = "kelvin", new_metric = "celsius"),
           rh = humidity::SH2RH(SPFH, TMP, p = PRES, isK = TRUE),
           rh = case_when(
             rh > 100 ~ 100,
             rh <= 100 ~ rh
           ),
           dp_c = humidity.to.dewpoint(rh = rh, t = tmp_c, temperature.metric = "celsius"),
           heat_index = weathermetrics::heat.index(t = tmp_f,
                           rh = rh,
                           temperature.metric = "fahrenheit",
                           output.metric = "fahrenheit",
                           round = 2))  


nldas_wide %>% 
  inspectdf::inspect_na()

nldas_wide_end2019 %>% summary()
```


## Wet Bulb Globe Temperature (WBGT) Estimate

We will apply a wrapper of the C code developed by Liljegren *et al* to estimate WBGT from datasets of standard meterological measurements . This product was produced by UChicago Argonne, LLC under contract with the Department of Energy [(GitHub link)](https://github.com/mdljts/wbgt). The source code is available at: [(link)](https://github.com/mdljts/wbgt/blob/master/src/wbgt.c).  
  
  
  
In addition to the weather variables and time, we will need `lat` and `lon` columns for the WBGT function. 

```{r lat_lon, eval = FALSE}

# Add lat/lon to nldas dataframe

  # Create lat lon table of centroids
  
base_coordinates <- 
  read_rds("data/select_bases.rds") %>%
  as_tibble() %>% 
  dplyr::select(site_name, centroid) %>%  
  mutate(longitude = unlist(map(centroid,1)),
         latitude = unlist(map(centroid,2)),
         site_name = stringr::str_replace_all(site_name, " ", "_"),
         site_name =  stringr::str_to_lower(site_name)) %>% 
  dplyr::select(-centroid)

base_coordinates 

```

WBGT with `mdljts` wrapper. Create a dataframe with the required columns for the `wbgt_df` function. 

```{r wbgt_mdljts, eval = FALSE}
# Create dataframe with columns needed for wbgt::wbgt_df (mdljts wrapper)

                     
# Join lat and lon
nldas_wide_end2019 <-
  nldas_wide_end2019 %>% 
  left_join(base_coordinates, by = c("installation" = "site_name")) %>% 
  dplyr::select(utc_dttm, local_dttm, installation, longitude, latitude, everything())

nldas_wide_end2019  

# Create datframe for `wbgt_df` function

wbgt_mdljts_df_end2019 <-
  nldas_wide_end2019 %>% 
    mutate(
      year = lubridate::year(local_dttm), 
      month = lubridate::month(local_dttm), 
      day = lubridate::day(local_dttm), 
      hour = lubridate::hour(local_dttm), 
      minute = lubridate::minute(local_dttm), 
      gmt = lubridate::hour(utc_dttm), 
      avg = 60, 
      lat = latitude, 
      lon = longitude, 
      solar = DSWRF, 
      pres = PRES / 100, 
      Tair = tmp_c, 
      relhum = rh, 
      speed = windspeed, 
      zspeed = 10, 
      dT = 0, 
      urban = 0) %>% 
  dplyr::select(year:urban)

wbgt_mdljts_df_end2019


# Execute WBGT wrapper on entire dataframe 
wbgt_mdljts_end2019 <-     
 wbgt::wbgt_df(wbgt_mdljts_df_end2019) %>% 
   dplyr::select(wbgt)


# Bind WBGT column to NLDAS dataframe and add Fahrenheit conversion

nldas_wide_end2019 <-
  nldas_wide_end2019 %>% 
  bind_cols(wbgt_mdljts_end2019) %>% 
  mutate(wbgt_f = weathermetrics::convert_temperature(wbgt, old_metric = "celsius", new_metric = "fahrenheit"))

# write_rds(nldas_wide_end2019, path = "data/nldas_wide_end2019.rds")

```

## "Wide" NLDAS dataframe

```{r open_nldas_wide, eval = FALSE}

# View "wide" NLDAS dataframe
# 
nldas_wide <-
  read_rds("data/nldas_wide.rds")

nldas_wide

nldas_wide_end2019 %>% summary

# Inspect for missing values

nldas_wide_end2019 %>% 
  inspectdf::inspect_na()


# Inspect date coverage

## Unique dates
nldas_wide_end2019 %>% 
  group_by(installation) %>% 
  summarize(unique_utc = n_distinct(utc_dttm)) %>% 
  arrange(unique_utc) 


nldas_wide %>% 
  group_by(installation) %>% 
  summarize(unique_local = n_distinct(local_dttm)) %>% 
  arrange(unique_local)

## Duplicate dates
dup_times <-
nldas_wide_end2019 %>% 
    group_by(installation) %>% 
    janitor::get_dupes(local_dttm)

dup_times %>% count(local_dttm)

  # Duplicates present in local time, occur at 1am local during Oct/Nov time change


# Mean temperature (K) from Jan 1990 - July 2019 (year-round)
nldas_wide %>% 
  group_by(installation) %>% 
  summarise(mean_tmp = mean(TMP)) %>% 
  arrange(desc(mean_tmp))

# Mean WBGT (F) from Jan 1990 - July 2019 (year-round)
nldas_wide %>% 
  group_by(installation) %>% 
  summarise(mean_wbgt_f = mean(wbgt_f)) %>% 
  arrange(desc(mean_wbgt_f))
```

## "Long" NLDAS dataframe

```{r convert_long_df, eval = FALSE}

# Gather by local time and location
# (90+ million rows for 25 locations, 1990-2018 hourly)


nldas_long_end2019 <-
  nldas_wide_end2019 %>%
    dplyr::select(installation, utc_dttm, tmp_f, rh, heat_index, wbgt_f) %>% 
    tidyr::gather(key = "variable", value = "value", tmp_f:wbgt_f) 
  


write_rds(nldas_long_end2019, path = "data/nldas_long_end2019.rds")


```

```{r view_nldas_long, eval = FALSE}

nldas_long <-
  read_rds("data/nldas_long.rds")

nldas_long_end2019
```



```{r combine_end2019, eval = FALSE}
nldas_wide <-
  read_rds("data/nldas_wide.rds")


nldas_long <-
  nldas_wide %>%
    dplyr::select(installation, utc_dttm, tmp_f, rh, heat_index, wbgt_f) %>% 
    tidyr::gather(key = "variable", value = "value", tmp_f:wbgt_f) 

nldas_long <-
  nldas_long %>% 
    bind_rows(nldas_long_end2019)


write_rds(nldas_long, path = "D:/nldas_long.rds")

nldas_wide <-
  nldas_wide %>% 
  bind_rows(nldas_wide_end2019)

nldas_wide

# write_rds(nldas_wide, path = "data/nldas_wide.rds")


```




## Nested NLDAS dataframe 
Create List-column for installation/variable pairs 

```{r nldas_nest, eval = FALSE}
# Increase memory limit
memory.limit()
memory.limit(size = 56000)

nldas_nest <-
  nldas_long %>% 
  dplyr::select(installation, variable, utc_dttm, value) %>% 
  nest(., data = utc_dttm:value)

# write_rds(nldas_nest, path = "data/nldas_nest.rds")

```

```{r view_nldas_nest, eval = FALSE}

nldas_nest <-
  read_rds("data/nldas_nest.rds")

nldas_nest

nldas_nest %>% 
  filter(installation == "eglin_afb" & variable == "tmp_f") %>% 
  pull(data)


```

