---
title: "extract_cells"
author: "SL"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/slewa/Projects/heat_stress")

library(sf)
library(tidyverse)
library(viridis)
library(rvest)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(lwgeom)
library(ncdf4)
library(devtools)
library(raster)
library(ncdump)
library(lubridate)
library(flipTime)
library(fuzzyjoin)
library(RNetCDF)
library(tidync)
library(furrr)
library(tabularaster)

```


# Extract variables from cells

This code extracts variables from NLDAS-2 netCDF files by converting .nc4 to raster and selecting by raster cell position. It includes a function to extract a single file and applies purrr:map2_dfc to map over each file and each defined variable.


## NLDAS-2 Data

List NLDAS-2 .nc4 files from 1990 - present.  
Extract date and hour from file name.

```{r extract_variable_names}

#file_list <-  
#  list.files(path = "D:/nldas", pattern = "^.*\\.(nc4|NC4|Nc4|Nc4)$",
#                        full.names = FALSE) %>% 
#  as_tibble() %>% 
#  mutate(date = value %>% 
#           stringr::str_extract("[1-2][0-9]{7}") %>% lubridate::ymd(),
#         hour = value %>% 
#           stringr::str_extract("[0-2][0-9][0]{2}") %>% as_factor(),
#         path = paste0("D:/nldas/", value)) %>% 
#  dplyr::select(value, path, date, hour) %>% 
#  filter(date >= "1990-01-01")

# saveRDS(file_list, "data/file_list.rds")

file_list <- readRDS("data/file_list.rds")
file_list

# Vector of NLDAS-2 file names

#file_names <-
#  file_list$value %>%
#  as_tibble() %>% 
#  dplyr::pull()


# Select NLDAS variables
nldas_variables <- c("TMP", "SPFH", "PRES", "UGRD", "VGRD", "DSWRF")

```

## Open ncdf4 files and read as Raster

Open a single .nc4 file and identify raster cell numbers from centroid coordinates.

```{r id_raster_cells, warning = FALSE}

# Load list of locations (object saved in shapefiles.Rmd)

select_bases <- readRDS("data/select_bases.rds")
select_bases

# Matrix of centroid X, Y coordinates

coordinates_matrix <- st_coordinates(select_bases$centroid)


# Open one .nc4 file as a raster and extract raster cellnumber (different from NLDAS grid ID)
# .nc4 files stored on external drive, not in project data folder due to size



# Open single raster file
  # Review dimensions (224 x 464), resolution (0.125 degree), extent, crs (WGS84) 
 
# single_raster <- raster(file_list$path[1], varname = "TMP", quick = "TRUE")

# saveRDS(single_raster, file = "data/single_raster.rds")

single_raster <- readRDS("data/single_raster.rds")
single_raster


#cells <- tabularaster::cellnumbers(single_raster, coordinates_matrix) %>% 
#    dplyr::select(cell_)

rm(single_raster)

# saveRDS(cells, file = "data/cells.rds")

cells <- readRDS("data/cells.rds")

# Join raster cells with site names

#cell_key <- cbind(select_bases$site_name, cells, stringsAsFactors = FALSE) %>% 
#  rename(site_name = `select_bases$site_name`)

# saveRDS(cell_key, file = "data/cell_key.rds")

cell_key <- readRDS("data/cell_key.rds")
cell_key

```

## Extract selected variables from selected cells

Function to extract selected variables from selected cells of a single NLDAS .nc4 file (unique GMT hour).  

Command to map single file function over list of files and list of variables (map2 as dataframe). This process may take approximately 40 hours of run time to process 29 years of data.


```{r extract_function_map, eval = FALSE}

# multiple variables
# 
nldas_path <- "D:/nldas/"

# output_path <- "D:/nldas/nldas_raster/"
output_path <- "D:/nldas/nldas_raster_joint/"


# Select cells non-Army bases (+ FSH):

cells <-
  cell_key %>% 
    filter(site_name %in% c(
           "MCB Quantico", "MCRD San Diego", "Twentynine Palms Main Base",
           "MCB Camp Lejeune", "MCRD Beaufort Parris Island", "MCB Camp Pendleton",
           "Eglin AFB", "Lackland AFB", "Fort Sam Houston")) %>% 
    dplyr::pull(cell_) 
    




# Function to extract selected variables from selected cells of a single NLDAS .nc4 file (unique GMT hour) 
  
extract_single_file <- function(ncdf_single_file, nldas_var) {  
  hourly_df <-   
  ncdf_single_file %>% 
      file.path(nldas_path, .) %>% 
        raster::brick(varname = nldas_var, quick = TRUE) %>%
        raster::extract(., cells, df = FALSE) %>% 
        magrittr::set_colnames(nldas_var) %>%
    as_tibble()
}

# Compile function
extract_single_file <- compiler::cmpfun(extract_single_file)




# Execute map over files

ptm <- proc.time()

for (i in seq_along(file_names)) {
map2_dfc(.x = file_names[[i]], .y = nldas_variables, .f = extract_single_file) %>% 
write_rds(., path = paste0("C:/Users/slewa/Documents/data/heat/nldas_raster_joint/",
                           paste0(
    file_names[[i]] %>%  stringr::str_extract("[1-2][0-9]{7}"),
    "_",
    file_names[[i]] %>% stringr::str_extract("\\.[0-2][0-9][0]{2}") %>% str_sub(start = 2L)),
    ".rds"), 
    compress = "none")
}

proc.time() - ptm



```

Single-file test of map command for an initial examination of output or for troubleshooting.  
```{r eval = FALSE}
# single instance test
map2_dfc(.x = file_names[[1]], .y = nldas_variables, .f = extract_single_file) %>% 
write_rds(., path = paste0("D:/nldas_raster_joint/",
                           paste0(
    file_names[[1]] %>%  stringr::str_extract("[1-2][0-9]{7}"),
    "_",
    file_names[[1]] %>% stringr::str_extract("\\.[0-2][0-9][0]{2}") %>% str_sub(start = 2L)),
    ".rds"), 
    compress = "none")

```


The map2 function above experienced malfunctions when reading through .nc4 files from an external hard drive. Errors would occur after minutes or hours of execution. 

The code below updates the file list to restart the map function.

```{r missing_files, eval = FALSE}
# Generate file list of missing files (for downloading in multiple iterations)

saved_list <- list.files(path = "C:/Users/slewa/Documents/data/heat/nldas_raster_joint/", pattern = "^.*\\.(rds)$",
                        full.names = FALSE) %>% 
  str_sub(end = -5L) %>% as_tibble()

file_names <-
  file_list %>% 
    mutate(day_hour = 
    paste0(
      file_list$value %>%  stringr::str_extract("[1-2][0-9]{7}"),
      "_",
      file_list$value %>% stringr::str_extract("\\.[0-2][0-9][0]{2}") %>% str_sub(start = 2L))) %>% 
    anti_join(., saved_list, by = c("day_hour" = "value")) %>% 
  dplyr::select(value) %>% 
  dplyr::pull()



```


## Compile hourly objects stacked in a list and save.

Combine individual objects (one for each hour) into a single master list.  

This step worked without issue for the first set of files (Army installations).  

The second set of files (other services) resulted in `readRDS` "unknown input format" errors, whether using `lapply` or `map` and regardless of whether the file directory was on an external drive or the computer hard drive.  
Troubleshooting: slice the files into sections. No issue with the first 100,000 files or second 100,000. Error between 200,000-258,901... tracked down to file #249,176: `"20180605_0700.rds"` (one bad file causing error). Deleted file and re-ran extract function.



```{r raster_list, eval = FALSE}
# https://stackoverflow.com/questions/6473831/readrdsfile-in-r/43509650#43509650

# setwd("C:/Users/slewa/Documents/data/heat/nldas_raster_joint")

nldas_list_joint_files <- list.files(pattern = ".rds") 

# read_rds(nldas_list_joint_files[249176])

nldas_list_joint <-
  map(nldas_list_joint_files, readRDS)

# write_rds(nldas_list_joint, path = "data/nldas_list_joint.rds")


## Troubleshooting: slice files, identify bad object
nldas_list_joint_1 <-
  nldas_list_joint_files %>%
    as_tibble() %>% 
    slice(1:100000) %>%
    pull(value) %>% 
    map(., readRDS)

nldas_list_joint_2 <-
  nldas_list_joint_files %>%
    as_tibble() %>% 
    slice(100001:200000) %>%
    pull(value) %>% 
    map(., readRDS)


  nldas_list_joint_files %>%
    as_tibble() %>% 
    slice(249175:249180) %>%
    pull(value) %>% 
    map(., readRDS)


```

## List structure
```{r examine_list}

nldas_list <- read_rds("data/nldas_list.rds")

# Review first element
nldas_list[1]

# Review list structure
nldas_list[1:3] %>% str



```

