---
title: "Shapefile and Grid Geometries"
author: "SL"
date: "July 1, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(sf)
library(tidyverse)
library(viridis)
library(rvest)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(lwgeom)
library(cowplot)
library(purrr)
library(purrrlyr)
library(tmap)
library(janitor)

```

## Load Military Installations, Ranges, and Training Areas (MIRTA) Dataset
Accessed from: https://catalog.data.gov/dataset/military-installations-ranges-and-training-areas. Metadata updated date: January 18, 2017
```{r load_shapefile}
#Ref: http://strimas.com/r/tidy-sf/
bases <- here::here("data/installations_ranges", "MIRTA_Boundaries.shp") %>%
  st_read() %>% 
  janitor::clean_names()

as_tibble(bases)
```


## Modify specific installations
```{r join_split_shapefiles}

## Join Fort Benning Georgia and Alabama shapefiles (st_union)
benning_join <-
  bases %>%
    filter(stringr::str_detect(site_name, "Fort Benning")) %>% 
    dplyr::select(geometry) %>% st_union() 


```

## Select installations
```{r select_sf_dataframe, warning = FALSE}
## Create dataframe of active duty Army installations, including FSH (JBSA) and single Fort Benning row 

army_bases <-
  bases %>%
    filter(oper_stat == "Active") %>% 
    mutate(component = replace(component, site_name == "Fort Sam Houston", "Army Active")) %>%
    filter(component == "Army Active") %>%
    filter(site_name != "Fort Benning") %>%
    mutate(geometry = replace(geometry, site_name == "Fort Benning GA", benning_geom ))
   

## Filter installations of interest

army_select <-
  army_bases %>% 
    dplyr::filter(.$site_name %in% c("Fort Benning GA", "Fort Bragg","Fort Campbell",
                              "Fort Jackson", "Fort Polk", "Fort Hood",
                              "Fort Stewart", "Fort Leonard Wood", "Fort Riley",
                              "NTC and Fort Irwin")) %>% 
    mutate(centroid = st_centroid(geometry)) %>% 
    dplyr::select(-c("component", "joint_base", "country", "oper_stat"))


st_crs(army_select)
as_tibble(army_select)
```

## Plot selected shapefiles

```{r plot_selected_bases}

## Plot selected installations
bases_plot <- purrr::map(army_select$site_name,
           function(x) {
             ggplot() +
               geom_sf(data = filter(army_select, site_name == x)) +
                ggtitle(x) +
                theme_bw() +
                theme(axis.text.x = element_text(size = rel(0.6)),
                      axis.text.y = element_text(size = rel(0.6))) 
           })

cowplot::plot_grid(plotlist = bases_plot)
```


```{r individual_base, include = FALSE, eval = FALSE}
bragg <-
  bases %>% 
    filter(.$site_name == "Fort Bragg")


# Save shapefile
# st_write(bragg, "bragg.shp")

st_centroid(bragg) %>% 
  as.tibble()

bragg_centroid <- st_centroid(bragg)

ggplot(bragg) +
  ggtitle("Fort Bragg") +
  geom_sf() +
  theme_bw()
```


## Load NLDAS grids
NLDAS grid shapefile from: https://ldas.gsfc.nasa.gov/sites/default/files/ldas/nldas/NLDAS_Grid_Reference.zip

```{r load_nldas_grid}

nldas_grid <- st_read("nldas_grids/NLDAS_Grid_Reference.shp") %>% 
  janitor::clean_names()

st_crs(nldas_grid)
as_tibble(nldas_grid)

```


## NLDAS and Installation Grid Overlap and Weighted Averages

```{r nldas_intersection_indiv, include = FALSE, eval = FALSE}
class(bragg)
bragg_nldas <-
  st_intersection(bragg, nldas_grid) %>% 
  mutate(area = sf::st_area(.),
         weight = area / sum(area))

bragg_intersects <-
  nldas_grid %>% filter(lengths(st_intersects(., bragg)) > 0)

as.data.frame(bragg_intersects) %>% 
  select(-geometry)

bragg_nldas

sum(bragg_nldas$area)
sum(bragg_nldas$weight)
st_area(bragg_nldas) 


ggplot(bragg_nldas) + 
  ggtitle("Fort Bragg NLDAS grids") +
  geom_sf() +
  geom_label(data = bragg_nldas, aes(x = centerx, y = centery, label = nldas_id), size = 3, fontface = "bold") +
  geom_text(data = bragg_nldas, aes(x = centerx, y = centery, label =  formatC(weight, format = "f", digits = 3)) , size = 3, position = position_nudge(y = -0.02)) +
  theme_bw()

```


```{r nldas_intersection_selected, warning = FALSE, message = FALSE}

# Identify all NLDAS grids intersected by installation shapefiles

bases_nldas = NULL
for (i in 1:nrow(army_select)) {
  base = army_select[i,] 
  base_nldas = st_intersection(base, nldas_grid) 
  bases_nldas = rbind(bases_nldas, base_nldas)
}
rm(base, base_nldas)
bases_nldas


# Calculate area of installation in each NLDAS grid and weight for spatial weighted average
nldas_weights <- bases_nldas %>% 
    mutate(area = sf::st_area(.$geometry)) %>% 
  group_by(site_name) %>% 
    mutate(sum_area = sum(area),
           weight = (area / sum_area)) %>% 
    dplyr::select(-c(nldas_x, nldas_y, area, centroid, sum_area))



# NLDAS grids that intersect with bases (to map full grid squares)
intersects = NULL
for (i in 1:nrow(army_select)) {
  base = army_select[i,] 
  base_intersect = nldas_grid %>% filter(lengths(st_intersects(., base)) > 0)
  intersects = rbind(intersects, base_intersect)
}
rm(base, base_intersect)

# Join `site_name` to selected NLDAS grid geometries

intersects <- intersects %>% left_join(dplyr::select(as_tibble(nldas_weights), c(nldas_id, site_name)), by = "nldas_id")
  
```


## Bounding box for installation
```{r bounding_box_indiv, eval=FALSE, warning = FALSE}
st_bbox(bragg) %>% 
  .[c("ymin", "xmin", "ymax", "xmax")]

bb_bragg <- st_as_sfc(st_bbox(bragg))
class(bb_bragg)


```


```{r bounding_box}

bbox = NULL
for (i in 1:nrow(army_select)) {
  base = army_select[i,] 
  base_bbox = st_bbox(base)
  bbox = rbind(bbox, base_bbox)
}
rm(base, base_bbox)
bbox

bbox_join <-
  as_tibble(army_select[,1]) %>% 
    bind_cols(as_tibble(bbox))

bbox_join %>% 
  dplyr::select(-geometry) %>% 
  print()

army_select <- bbox_join %>% 
  dplyr::select(-geometry) %>%
  right_join(army_select, by = "site_name")



```


## Plots of intersecting NLDAS grids
```{r grid_plots_indiv, eval = FALSE}
ggplot() + 
  geom_sf(data = bragg_nldas) +
  geom_sf(data = bb_bragg , color = "blue", fill = "NA") +
  geom_sf(data = bragg_centroid, color = "red") +
  ggtitle("Fort Bragg Bounding Box") +
  geom_label(data = bragg_nldas, aes(x = centerx, y = centery, label = nldas_id), size = 3, fontface = "bold") +
  geom_text(data = bragg_nldas, aes(x = centerx, y = centery, label =  formatC(weight, format = "f", digits = 3)) , size = 3, position = position_nudge(y = -0.02)) +
  theme_bw()


ggplot() +
  geom_sf(data = bragg_intersects) +
  geom_sf(data = bb_bragg , color = "blue", fill = "NA") +
  geom_sf(data = bragg_centroid, color = "red") +
  geom_sf(data = bragg, fill = "NA") +
  ggtitle("Fort Bragg Intersecting Grids") +
  geom_label(data = bragg_nldas, aes(x = centerx, y = centery, label = nldas_id), size = 3, fontface = "bold") +
  geom_text(data = bragg_nldas, aes(x = centerx, y = centery, label =  formatC(weight, format = "f", digits = 3)) , size = 3, position = position_nudge(y = -0.03)) +
  theme_bw()

```

```{r grid_plots}

# Faceted maps
tm_shape(intersects) +
  tm_borders() +
  tm_text("nldas_id", size = 0.5) +
  tm_shape(nldas_weights) +
  tm_borders() +
  tm_graticules(col = "gray90", alpha = 0.5, labels.size = 0.5) +
  tm_facets(by = "site_name") +
  tm_layout(panel.label.size = 0.8, panel.label.height = 0.8)


# tm_shape(intersects %>%  filter(site_name == "Fort Polk")) +
#  tm_borders() +
#  tm_text("nldas_id", size = 0.5) +
#  tm_shape(nldas_weights %>% filter(site_name == "Fort Polk")) +
#  tm_borders() +
#  tm_graticules(col = "gray90", alpha = 0.5, labels.size = 0.5) +
#    tm_layout(title = "Fort Polk")

 
# Individual site maps (tmap) 

site_list <- unique(as.character(army_select$site_name))

for (i in seq_along(site_list)) { 
site_plot_tmap <-
  tm_shape(intersects %>%  filter(site_name == site_list[i])) +
  tm_borders() +
  tm_text("nldas_id", size = 0.7) +
  tm_shape(nldas_weights %>% filter(site_name == site_list[i])) +
  tm_borders() +
  tm_fill(col = "darkolivegreen4", alpha = 0.2) +
  tm_graticules(col = "gray90", alpha = 0.3, labels.size = 0.8) +
    tm_layout(title = site_list[i])
print(site_plot_tmap)
}



# Individual site maps (ggplot) 

for (i in seq_along(site_list)) { 
site_plot_gg <-

ggplot() +
  geom_sf(data = intersects %>% filter(site_name == site_list[i])) +
  geom_rect(data = army_select %>% filter(site_name == site_list[i]), 
            aes(xmin = xmin , xmax = xmax, ymin = ymin, ymax = ymax), color = "blue", fill = "transparent") +
  ggtitle(site_list[i]) +
  geom_sf(data = nldas_weights %>% filter(site_name == site_list[i]), fill = "NA") +
  geom_label(data = nldas_weights %>% filter(site_name == site_list[i]), aes(x = centerx, y = centery, label = nldas_id), size = 3, fontface = "bold") +
  geom_text(data = nldas_weights %>% filter(site_name == site_list[i]), aes(x = centerx, y = centery, label =  formatC(weight, format = "f", digits = 3)) , size = 3, position = position_nudge(y = -0.03)) +
  theme_bw() +
  theme(axis.title = element_text())

print(site_plot_gg)
}

```
