---
title: "Shapefile and Grid Geometries"
author: "SL"
output:
  html_document:
    toc: true
    toc_float: true
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(sf)
library(tidyverse)
library(viridis)
library(rvest)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(lwgeom)
library(cowplot)
library(purrr)
library(purrrlyr)
library(tmap)
library(janitor)
library(plotly)

```

## Load Military Installations, Ranges, and Training Areas (MIRTA) Dataset

Accessed from: https://catalog.data.gov/dataset/military-installations-ranges-and-training-areas. Metadata updated date: January 18, 2017.

Join Fort Benning Georgia and Alabama shapefiles.

Fort Lewis (JBLM) geometry includes remote sites.

```{r load_shapefile}

bases <- here::here("data/installations_ranges", "MIRTA_Boundaries.shp") %>%
  sf::st_read(stringsAsFactors = FALSE) %>% 
  janitor::clean_names()

## Edit Fort Benning geometries
benning_join <-
  bases %>%
    filter(stringr::str_detect(site_name, "Fort Benning")) %>% 
    dplyr::select(geometry) %>% st_union() 

bases <- bases %>% 
  filter(site_name != "Fort Benning") %>%
  mutate(geometry = replace(geometry, site_name == "Fort Benning GA", benning_join ))

rm(benning_join)

## Edit Fort Lewis (JBLM) geometries to remove remote polygons
## Include sites near Yakima, WA and Camp Rilea, OR

# "Before" plot
bases %>% 
    filter(site_name == "Fort Lewis") %>% 
    ggplot() +
       geom_sf() +
       ggtitle("Fort Lewis (JBLM) - Before") 

# Set bounding box
jblm_filt_bbox <- 
  sf::st_bbox(c(xmin = -123, 
                ymin = 46.8, 
                xmax = -122, 
                ymax = 47.2), 
                crs = st_crs(4326)) %>% 
 sf::st_as_sfc(.)


# separate MULTIPOLYGON to POLYGON(s)

jblm_polygon <-
   bases %>% 
    filter(site_name == "Fort Lewis") %>% 
    sf::st_cast(., "POLYGON")

# identify polygons within bbox
jblm_new_geom <-
  sf::st_within(jblm_polygon$geometry, jblm_filt_bbox) 

# add column for within box status
jblm_polygon <-
  jblm_polygon %>% 
    mutate(within_bbox = lapply(jblm_new_geom, `[`, 1)) %>% 
    filter(within_bbox == 1) %>% 
    dplyr::select(-within_bbox) %>%
    sf::st_union()


# replace Fort Lewis geometry in main dataframe

bases <- bases %>% 
  mutate(geometry = replace(geometry, site_name == "Fort Lewis", jblm_polygon))

rm(jblm_filt_bbox, jblm_new_geom, jblm_polygon)

# "After" plot
bases %>% 
    filter(site_name == "Fort Lewis") %>% 
    ggplot() +
       geom_sf() + 
       ggtitle("Fort Lewis (JBLM) - After")    
```


## Select installations

Select bases for analyses. Citeria: in list of top CONUS bases for heat illnesses from MSMR report (April 2018). Additional Army installations based on HSI counts and rates (from DMED) and for diversity of climate regions.

Add centroid column.
```{r select_sf_dataframe, warning = FALSE}

# Look-up base name in MITRA list  
#  search_names <- bases$site_name %>% as.character()

#  stringr::str_subset(search_names, "Lewis")

select_bases <-
  bases %>%
    dplyr::filter(.$site_name %in% c("Fort Benning GA", "Fort Bragg","Fort Campbell",
                              "Fort Jackson", "Fort Polk", "Fort Hood",
                              "Fort Stewart", "Fort Leonard Wood", "Fort Riley",
                              "NTC and Fort Irwin", "Fort Bliss", "Fort Lewis",
                              "Fort Sill", "Fort Carson", "Fort Gordon",
                              "Fort Drum", "MCB Camp Lejeune", "MCRD Beaufort Parris Island",
                              "MCB Camp Pendleton", "MCRD San Diego", "MCB Quantico",
                              "Twentynine Palms Main Base", "Lackland AFB", "Eglin AFB", "Fort Sam Houston")) %>% 
    mutate(centroid = st_centroid(geometry)) %>% 
    dplyr::select(-c("joint_base", "country", "oper_stat")) 


saveRDS(select_bases, "data/select_bases.rds")

select_bases
```



```{r plot_selected_bases, eval = FALSE}

## Plot selected installations
purrr::map(select_bases$site_name,
           function(x) {
             ggplot() +
               geom_sf(data = filter(select_bases, site_name == x)) +
               geom_sf(data = st_centroid(filter(select_bases, site_name == x))) +
                ggtitle(x) +
                theme_bw() +
                theme(axis.text.x = element_text(size = rel(1)),
                      axis.text.y = element_text(size = rel(1))) 
           })

#cowplot::plot_grid(plotlist = bases_plot)
```


## Load NLDAS grids

NLDAS grid shapefile from: https://ldas.gsfc.nasa.gov/sites/default/files/ldas/nldas/NLDAS_Grid_Reference.zip

```{r load_nldas_grid}

nldas_grid <- 
  here::here("data/nldas_grids", "NLDAS_Grid_Reference.shp") %>%
  sf::st_read(stringsAsFactors = FALSE) %>% 
  janitor::clean_names()

nldas_grid

```


## NLDAS and Installation Grid Overlap 


```{r nldas_intersection_selected, warning = FALSE, message = FALSE}

# Identify all NLDAS grids intersected by installation shapefiles

bases_nldas = NULL

  for (i in 1:nrow(select_bases)) {
    i_base = select_bases[i,] 
    base_nldas = st_intersection(i_base, nldas_grid) 
    bases_nldas = rbind(bases_nldas, base_nldas)
  }

bases_nldas

rm(i_base, base_nldas)


# Calculate area of installation in each NLDAS grid and weight for spatial weighted average
nldas_weights <- bases_nldas %>% 
    mutate(area = sf::st_area(.$geometry)) %>% 
  group_by(site_name) %>% 
    mutate(sum_area = sum(area),
           weight = (area / sum_area)) %>% 
    dplyr::select(-c(nldas_x, nldas_y, area, centroid, sum_area))



# NLDAS grids that intersect with bases (to map full grid squares)
intersects = NULL

  for (i in 1:nrow(select_bases)) {
    i_base = select_bases[i,] 
    base_intersect = nldas_grid %>% filter(lengths(st_intersects(., i_base)) > 0)
    intersects = rbind(intersects, base_intersect)
  }

rm(i_base, base_intersect)

# Join `site_name` to selected NLDAS grid geometries

intersects <- intersects %>% left_join(dplyr::select(as_tibble(nldas_weights), c(nldas_id, site_name)), by = "nldas_id")

intersects  
```


## Plots with intersecting NLDAS grids
```{r grid_plots, warning = FALSE}

select_bases$site_name <- purrr::set_names(select_bases$site_name)


base_grid_plot_function <- function(x) {
      ggplot() +
                 geom_sf(data = filter(select_bases, site_name == x)) +
                 geom_sf(data = filter(intersects, site_name == x), fill = NA, color = "blue") +
                 geom_sf(data = st_centroid(filter(select_bases, site_name == x)), shape = 3, size = 3, stroke = 1.5) +
                 geom_label(data = filter(intersects, site_name == x), aes(x = centerx, y = centery, label = nldas_id), size = 2) +
                  ggtitle(x) +
                  theme_bw() +
                  theme(axis.text.x = element_text(size = rel(1)),
                        axis.text.y = element_text(size = rel(1)),
                        axis.title.x = element_blank(),
                        axis.title.y = element_blank()) 
}


# base_grid_plot_function(select_bases$site_name[1])


base_grid_plots <-
  map(select_bases$site_name, ~base_grid_plot_function(.x))

base_grid_plots
```



